---
title: "Pharmacotherapy for Pregnant Women with Opioid Use Disorder "
author: "<br>Ahra Kim, MPH<br>"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
---

<style type="text/css">
body, td{
  font-size: 14px;
}
code.r{
  font-size: 12px;
}
pre{
  font-size: 12px
}
div.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE, warning=FALSE, message=FALSE, comment=NA}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	comment = NA,
	cache = FALSE
)
setwd("/Volumes/p50_oud/Ahra") 

# suppress scientific notation
options(scipen = 999)
options(knitr.kable.NA = '')

library(Hmisc)
library(tangram)
library(readr)
library(haven)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(ggh4x)
# library(plotly)
# library(data.table)
library(reshape2)
library(janitor)
library(kableExtra)
# library(UpSetR)
library(survival)
library(rms)
library(mice)
library(PSW)


# source("make_cph_table.R")
options(prType="html")

# table style to show percent complete/missing N in demographics
nejm <- tangram::nejm # Copy style/transform into local environment

# Override the generation of N values for the transform
nejm[["Cell"]][["n"]] <- function(n, class=NULL, hdr=FALSE, possible=NULL, ...)
{
  n <- if(hdr) paste0("(N=",as.character(n),")") else
       if(is.null(possible)) as.character(n) else
         paste0(as.character(n), " (", render_f(as.numeric(n)/possible*100, 2), ")")
  
  cell(n, class=c("cell_n", class), ...)
}


# load("/Volumes/p50_oud/Ahra/G5_data_202503.Rdata")
# load("/Volumes/p50_oud/Ahra/G5_data_202506.Rdata")
load("/Volumes/p50_oud/Ahra/G5_data_202509.Rdata")
ndc <- read.csv("/Volumes/p50_oud/Ahra/NDCs_2020.csv", colClasses = c(ndcnum = "character"))
# ndctest <- ndc %>% group_by(prodnme, gennme) %>%  summarise(ndcnum = paste(unique(ndcnum), collapse = ", "),
#                                                             .groups = 'drop') # for manuscript
# 
# write.csv(ndctest, file="ndctest.csv", row.names = FALSE)

# deathcode <- read_delim("~/Documents/Research/IDP-Spring2023/ValueSet_PHVS_CauseOfDeath_ICD-10_CDC_V2_20230811-114942/ValueSets/PHVS_CauseOfDeath_ICD-10_CDC_V2.txt", delim = "\t", escape_double = FALSE, trim_ws = TRUE)
# deathcode <- deathcode[4:10853, ]

# filter people with the drug of interests: buprenorphine, naltrexone, methadone
# Per Ashley add extra codes (July 2024)
# 
# # most of these medications no longer in new data
extract <- af_extract[which(af_extract$ndc %in% c(ndc$ndcnum, 'Q9991',
                                                  'Q9992',
                                                  'J0571',
                                                  'J0572',
                                                  'J0573',
                                                  'J0574',
                                                  'J0575',
                                                  'G2068',
                                                  'G2069',
                                                  'G2070',
                                                  'G2072',
                                                  'G2079',
                                                  'J0575',
                                                  'G2068',
                                                  'G2069',
                                                  'G2070',
                                                  'G2079',
                                                  'G2072',
                                                  'J2315',
                                                  'G2073',
                                                  'G2057',
                                                  'G2067',
                                                  'G2078',
                                                  'S0109')), ]
methadonereceipt <- af_extract[af_extract$ndc %in% c("H0020", "G2067", "G2078"), ]
extract <- extract[!extract$StudyID %in% methadonereceipt$StudyID, ]


af_extract <- as.data.frame(af_extract)
extract <- af_extract[af_extract$ndc %in% ndc$ndcnum, ]

# test <- table(af_extract$ndc)
# tail(test, 20)
# testH0020 <- af_extract[af_extract$ndc %in% "H0020",]
# j1230 <- as.data.frame(af_extract[af_extract$ndc == "J1230", ]) # they are already in the af_extract table

extract <- merge(extract, ndc, by.x="ndc", by.y= "ndcnum", all.x = TRUE)
extract <- extract[extract$BTrack==90, ]
extract$gennme <- ifelse(is.na(extract$gennme) & extract$ndc =="H0020", "Methadone", extract$gennme)
extract$daysup <- ifelse(is.na(extract$daysup) & extract$ndc =="H0020", 1, extract$daysup)

# subset people who are in the exposure group
# basic <- af_master_new[af_master_new$StudyID %in% extract$StudyID, ]
# basic <- af_master[af_master$Expose1==1, ] # 998421 to 31199
basic <- af_master
# exclude people if they have an overdose or death from -90 to 41 days
# not a problem in the new data; everyone has outcome of 42+ days
# af_outcome$dead_before_follow <- ifelse(!is.na(af_outcome$death_day) & af_outcome$death_day <=41, 1, 
#                              ifelse(!is.na(af_outcome$overdose_day) & af_outcome$overdose_day <= 41, 1, 0))
# dead_before_follow <- af_outcome[af_outcome$dead_before_follow==1, ]
# af_outcome <- af_outcome[af_outcome$dead_before_follow==0 , ]

# want 90 day only since 180 is for sensitivity
basic <- basic[basic$BTrack==90, ] # n=15951
basic$filledrx <- ifelse(basic$StudyID %in% extract$StudyID, 1, 0)

af_master$unique_id <- sub(".*-", "", af_master$StudyID)
basic$unique_id <- sub(".*-", "", basic$StudyID)

# merge covariates
af_cov$unique_id <- sub(".*-", "", af_cov$StudyID)
# only include one of the pregnancies if twins/triplets 
basic <- basic[order(basic$unique_id, basic$StudyID), ]
# separate those with multiple pregnancies
# create indicators: if next row covariates are the same as previous row then they are twins/triplets
basic <- basic %>% mutate(rownum= row_number()) %>% data.frame()
# possibly twins? 
# these are the records where the study id is duplicated (either twins or multiple children over years)
twins <- basic %>% group_by(unique_id) %>% 
            mutate(n = n()) %>% # count number of same study id
                filter(n> 1) %>%
                        mutate(mom_dob_prevrow = lag(MomDOB_day), 
                            gweeks20_prevrow = lag(GWeeks20_dt_day))
twins$multbirths <- ifelse(twins$mom_dob_prevrow == twins$MomDOB_day & twins$gweeks20_prevrow == twins$GWeeks20_dt_day, 1, 0)
twins$multbirths <- ifelse(is.na(twins$multbirths), 0, twins$multbirths)  
not_twins <- twins[twins$multbirths==0 ,]
twins2 <- twins[twins$multbirths==1,] # these are twins/triplets

# exclude twins/triplets
basic <- basic[! (basic$rownum %in% twins2$rownum),  ] # n=15328


# some covariates are in Btrack 180 only and not 90
af_cov <- as.data.frame(af_cov)
# af_cov2 <- af_cov[af_cov$BTrack==90, ]
af_cov2 <- af_cov[af_cov$StudyID %in% basic$StudyID, ]

af_cov_u <- af_cov2[ , !(colnames(af_cov2) %in% c("BTrack", "condition", "condition_from_day", "CodeID"))]
af_cov_u <- af_cov_u[!duplicated(af_cov_u), ]

basic <- merge(basic, af_cov_u, by= "StudyID", all.x = TRUE)

# df <- basic #creating a copy to work off of just in case
names(basic) <- tolower(names(basic))
```

```{r}

# duplicated ids 
# dups <- basic[which(duplicated(basic$unique_id)), "unique_id"]
# dup <- basic[which(duplicated(basic$unique_id)), ]
# dupdf <- basic[which(basic$unique_id %in% dups), ]
# include only the first pregnancy for now 

# Format covariates
label(basic$momage) <- "Age at delivery"
basic$momrace <- ifelse(basic$momrace=="", NA, basic$momrace)
basic$momrace_desc <- ifelse(basic$momrace_desc=="", NA, basic$momrace_desc)
basic$momethnicity_desc <- ifelse(basic$momethnicity_desc=="", NA, basic$momethnicity_desc)
basic$momrace.new <- ifelse(basic$momrace_desc %in% c("White", "Bridged White"), "White",
                            ifelse(basic$momrace_desc %in% c("Black or African-American", "Bridged Black"), "Black of African American",
                                   ifelse(basic$momrace_desc %in% c("American Indian or Alaskan Native", "Bridged American Indian or Alaskan Native"), "American Indian or Alaska Native",
                                      ifelse(basic$momrace_desc %in% c("Asian Indian", "Bridged Asian or Pacific Isander", "Chinese", "Filipino", "Korean", "Native Hawaiian", "Other Asian", "Vietnamese"), "Asian or Pacific Islander",
                                             ifelse(basic$momrace_desc %in% c("Other Race", "Refused or Unknown"), "Other or Unknown", NA)))))

label(basic$momrace.new) <- "Race"

basic$momhisp.new <- ifelse(basic$momethnicity_desc %in% c("\tNot Hispanic", "\tOther", "\tUnknown"), "Not Hispanic", 
                            ifelse(is.na(basic$momethnicity_desc), NA, "Hispanic"))

basic$momraceeth.new <- ifelse(basic$momhisp.new=="Hispanic", "Hispanic",
                               ifelse((basic$momhisp.new!="Hispanic"| is.na(basic$momhisp.new)) & basic$momrace.new=="Asian or Pacific Islander", "Non-Hispanic Asian",
                                      ifelse((basic$momhisp.new!="Hispanic"| is.na(basic$momhisp.new)) & basic$momrace.new=="Black of African American", "Non-Hispanic Black",
                                             ifelse((basic$momhisp.new!="Hispanic"| is.na(basic$momhisp.new)) & basic$momrace.new=="White", "Non-Hispanic White", NA))))
basic$momraceeth.new <- ifelse(basic$momhisp.new=="Not Hispanic" & is.na(basic$momrace.new) & is.na(basic$momraceeth.new), "Non-Hispanic Other", 
                               ifelse(is.na(basic$momhisp.new) & basic$momrace.new=="White" & is.na(basic$momraceeth.new), "Non-Hispanic White",
                                      ifelse(is.na(basic$momhisp.new) & basic$momrace.new %in% c("American Indian or Alaska Native", "Other or Unknown"), "Non-Hispanic Other", 
                                             ifelse(basic$momhisp.new=="Not Hispanic" & basic$momrace.new %in% c("American Indian or Alaska Native", "Other or Unknown") & is.na(basic$momraceeth.new), "Non-Hispanic Other", basic$momraceeth.new))))
label(basic$momraceeth.new) <- "Race/Ethnicity combined"

label(basic$rpl_themes) <- "CDC SVI overall percentile ranking of all 4 themes"


#  information related to each delivery/pregnancy
basic$routedeliv.new <- ifelse(basic$deliverycesarean=="Y", "Cesarean", ifelse(basic$deliveryvaginal=="Y", "Vaginal", ifelse(basic$deliveryforcep=="Y", "Vaginal", ifelse(basic$deliveryvacuum=="Y", "Vaginal", NA))))
basic$routedeliv.new <- as.factor(basic$routedeliv.new)
label(basic$routedeliv.new) <- "Route of delivery"

# exclude if route of delivery everything is marked as "NO"
# excluded one person
basic <- basic[!is.na(basic$routedeliv.new), ] #15328 to 15272

label(basic$previouspregnanciestotaln) <- "Number of previous pregnancies"
label(basic$cat4_seg1) <- "Enrollment up to 365 days post-partum"

basic$enroll365.new.f <- factor(basic$cat4_seg1, levels = c(0,1), labels=c("No", "Yes"))
label(basic$enroll365.new.f) <- "Enrollment up to 365 days post-partum"

label(basic$momhisp.new) <- "Hispanic"

# Death
# want to capture deaths from day 42 and onwards but before loss of enrollment; some deaths captured after they disenroll (for now, but will go back and capture them later)
basic$death.new <- ifelse(!is.na(basic$momdeath_day) & basic$momdeath_day <= 323 & basic$momdeath_day <= basic$fdate_seg1_day, 1, 0)
basic$death.new.f <- factor(basic$death.new, levels=c(0, 1), labels=c("No", "Yes"))
label(basic$death.new.f) <- "Death during follow-up only (not after disenroll)"

basic$death.new.2 <- ifelse(!is.na(basic$momdeath_day) & basic$momdeath_day <= 323 , 1, 0)
basic$death.new.2.f <- factor(basic$death.new.2, levels=c(0, 1), labels=c("No", "Yes"))
label(basic$death.new.f) <- "Death within 1 year (even after disenroll)"

# to check deaths
# dead323 <- basic[which(basic$cat2_seg1==1 & basic$momdeath_day<= 323), ]
```


```{r}
# comorbidities
af_cov2 <- af_cov2[af_cov2$condition!="", ]
af_cov2 <- af_cov2[af_cov2$BTrack==90, ]

# transform conditions from long to wide
af_cov3 <- af_cov2[ , c("StudyID", "condition_from_day", "condition")]

# comorb <- af_cov3 %>% group_by(StudyID) %>%
#   pivot_wider(id_cols = "StudyID", names_from = "condition", values_from = "condition") %>%
#   mutate_all(~replace(., . == "NULL", 0)) %>%
#   mutate_all(~replace(., . != "0", 1)) %>% 
#   mutate_if(is.list, as.numeric) %>%
#   ungroup() %>% as.data.frame()

comorb <- af_cov3 %>% group_by(StudyID) %>%
  pivot_wider(id_cols = "StudyID", names_from = "condition", values_from = "condition") %>%
  mutate(across(everything(), ~replace(., . == "NULL", 0))) %>%
  mutate(across(everything(), ~replace(., . != "0", 1))) %>% 
  mutate_if(is.list, as.numeric) %>%
  ungroup() %>% as.data.frame()

basic <- merge(basic, comorb, by.x = "studyid", by.y = "StudyID", all.x = TRUE)

# remove those with missing comorbidities
basic <- basic[basic$studyid %in% comorb$StudyID, ]

comorbidities <- names(comorb)[2:18]
comorblabs <- gsub("_", " ", comorbidities)
comorblabs <- capitalize(comorblabs)
comorblabs <- gsub("nsaids", "NSAIDS", comorblabs)
comorblabs <- gsub("substance", "substance use", comorblabs)
comorblabs <- gsub("Bipolard", "Bipolar disorder", comorblabs)
comorblabs <- gsub("psych", "psychotic disorders", comorblabs)
comorblabs <- gsub("benzo", "benzodiazepine use", comorblabs)
comorblabs <- gsub("usedis", "use disorder", comorblabs)
comorblabs <- gsub("anticonv", "anticonvulsant use", comorblabs)
comorblabs <- gsub("Personality dis", "Personality disorder", comorblabs)
comorblabs <- gsub("Alc", "Alcohol", comorblabs)
comorblabs <- gsub("Mat", "Maternal", comorblabs)
comorblabs <- gsub("relax", "relaxant use", comorblabs)
comorblabs <- gsub("Med barbit", "Med barbiturates", comorblabs)

# barbiturates are new in data as of January 2026 after removing Exposure1 criteria


# create factor variables
for(i in 1:length(comorbidities)) {
  basic[[paste0(comorbidities[i], ".f")]] <- basic[[comorbidities[i]]]
  basic[[paste0(comorbidities[i], ".f")]] <- ifelse(is.na(basic[[paste0(comorbidities[i], ".f")]]), 0, basic[[paste0(comorbidities[i], ".f")]])
  basic[[paste0(comorbidities[i], ".f")]] <- factor(basic[[paste0(comorbidities[i], ".f")]], levels = c(0:1), labels=c('No', 'Yes'))
  label(basic[[paste0(comorbidities[i], ".f")]]) <- comorblabs[i]
}

label(basic$gweeks) <- "Number of Gestation weeks"

# Combine variables for analyses
basic$pain.combine <- ifelse(basic$back_pain.f=="Yes" | basic$dental_pain.f=="Yes" | basic$fibromyalgia.f=="Yes" | basic$med_nsaids.f=="Yes" | basic$med_muscle_relax.f=="Yes", 1, 0)
basic$pain.combine.f <- factor(basic$pain.combine, levels = 0:1, labels=c("No", "Yes"))
label(basic$pain.combine.f) <- "Combined pain"

basic$personality.combine <- ifelse(basic$personality_dis.f=="Yes"| basic$schizophrenia_psych.f=="Yes", 1, 0)
basic$personality.combine.f <- factor(basic$personality.combine, levels = 0:1, labels=c("No", "Yes"))
label(basic$personality.combine.f) <- "Schizophrenia or personality disorder"

basic$alcuse.othersub <- ifelse(basic$alc_usedis.f=="Yes" | basic$other_substance.f=="Yes", 1, 0)
basic$alcuse.othersub.f <- factor(basic$alcuse.othersub, levels = 0:1, labels=c("No", "Yes"))
label(basic$alcuse.othersub.f) <- "Alcohol or other substance use"

basic$benzo.anticonv <- ifelse(basic$med_benzo.f=="Yes" | basic$med_anticonv.f == "Yes", 1, 0)
basic$benzo.anticonv.f <- factor(basic$benzo.anticonv, levels=0:1, labels=c("No", "Yes"))
label(basic$benzo.anticonv.f) <- "Benzodiazepine or anticonvulsant use"

basic$anxiety <- ifelse(is.na(basic$anxiety), 0, basic$anxiety)
basic$depression <- ifelse(is.na(basic$depression), 0, basic$depression)
basic$mat_morbidity <- ifelse(is.na(basic$mat_morbidity), 0, basic$mat_morbidity)


```

```{r}
# overdose
af_out <- af_outcome[af_outcome$BTrack==90 & af_outcome$StudyID %in% basic$studyid, ]
names(af_out) <- tolower(names(af_out))
# change to count outcomes that occur after disenrollment too
af_out$fdate_seg1_day <- basic[match(af_out$studyid, basic$studyid), "fdate_seg1_day"]
# check to see which overdose or death occurs after disenrollment
# x <- af_out[which(af_out$overdose_day - 42 > af_out$fdate_seg1_day), ]
# subset events that occured before disenrollment (before)
af_out$momdeath_day <- basic[match(af_out$studyid, basic$studyid), "momdeath_day"]
# af_out <- af_out[(!is.na(af_out$overdose_day) & (af_out$overdose_day <= af_out$fdate_seg1_day)) | (!is.na(af_out$momdeath_day) & (af_out$momdeath_day<= af_out$fdate_seg1_day)), ]
af_out$momdeath_day <- ifelse(af_out$momdeath_day > 323, NA, af_out$momdeath_day)

# outcomes file contains multiple overdose records (previous ODs can show up on next visits)
# summarize information on total number of overdoses
af_out2 <- af_out %>% group_by(studyid) %>%
  filter(overdose_opioid_outcome ==1  | overdose_nonopioid_outcome ==1) %>%
  mutate(n= n()) %>% data.frame()

af_out2$unique_id <- sub(".*-", "", af_out2$studyid)

# exclude if they had OD in both exposure and outcome
af_out2 <- af_out2[!(af_out2$overdose_nonopioid_exposure==1 & af_out2$overdose_nonopioid_outcome==1), ]
af_out2 <- af_out2[!(af_out2$overdose_opioid_exposure==1 & af_out2$overdose_opioid_outcome==1), ]
af_out2 <- af_out2[!(af_out2$overdose_nonopioid_exposure==1 & af_out2$overdose_opioid_outcome==1), ]
af_out2 <- af_out2[!(af_out2$overdose_opioid_exposure==1 & af_out2$overdose_nonopioid_outcome==1), ]

af_out2 <- af_out2[af_out2$overdose_nonopiod_day >= 0 & af_out2$overdose_nonopiod_day <= 323 | 
                     af_out2$overdose_opiod_day >= 0 & af_out2$overdose_opiod_day <= 323,  ]

# If someone who had multiple births had overdoses for both births, then both information will be included (but this did not happen in the data)



# af_out2 <- af_out %>% group_by(studyid) %>%
#   filter(!is.na(overdose_day)) %>%
# #   mutate(n= n()) %>% data.frame()
# 
# basic$num.od <- af_out2[match(basic$studyid, af_out2$studyid), "n"]
# label(basic$num.od) <- "Number of any overdoses recorded during follow-up"

# Filter out opioid related overdose only
# af_out_op <- af_out %>% group_by(studyid) %>%
#   filter(overdose_opioid_outcome==1) %>%
#   mutate(n= n()) %>% data.frame()
# 
# basic$num.opioid.od <- af_out_op[match(basic$studyid, af_out_op$studyid), "n"]
# label(basic$num.opioid.od) <- "Number of opioid overdoses recorded during follow-up"  

# ever had od during follow-up
# basic$outcome.od <- ifelse(!is.na(basic$num.od), 1, 0)
# label(basic$outcome.od) <- "Had any overdose during follow-up"

# ever had opioid od during follow-up
# basic$outcome.opioid.od <- ifelse(!is.na(basic$num.opioid.od), 1, 0)
# label(basic$outcome.opioid.od) <- "Had any opioid overdose during follow-up"
#   

# More deaths from outcome table; merge into large table
af_out_death <- af_out %>% group_by(studyid) %>% 
  filter(!is.na(momdeath_day) & momdeath_day <= 323) %>% data.frame()

af_out_death <- af_out_death[ , c("studyid", grep("dcause", names(af_out_death), value=TRUE), "momdeath_day")]
af_out_death <- unique(af_out_death)
# basic <- merge(basic, af_out_death, by="studyid", all.x = TRUE)
basic$death_day <- basic$momdeath_day
basic$death_day <- ifelse(basic$death_day > 323, NA, basic$death_day)

```


```{r}
# get dosage information
# redbook <- read_sas("/Volumes/HPDrdw/MarketScan/RawData/CCAE/Random_1Pct/SAS Data 2022/redbook_yr22.sas7bdat", NULL)
# names(redbook) <- tolower(names(redbook))
# redbook$gennme2 <- tolower(redbook$gennme)
# redbook$methadone <- ifelse(grepl("methadone", redbook$gennme2), 1, 0)
# redbook_methadone <- redbook[redbook$methadone==1, ]
# write.csv(redbook_methadone, file="redbook_methadone.csv", row.names = FALSE)
redbook_methadone <- read.csv("/Volumes/p50_oud/Ahra/redbook_methadone.csv")
redbook_methadone$ndcnum  <-sprintf("%0*s", 11, as.character(redbook_methadone$ndcnum)) 

# redbook2 <- redbook[redbook$ndcnum %in% extract$ndc, ]
# write.csv(redbook2, file="redbook2.csv", row.names = FALSE)
redbook2 <- read.csv("/Volumes/p50_oud/Ahra/redbook2.csv")
redbook2$ndcnum <-sprintf("%0*s", 11, as.character(redbook2$ndcnum)) 
extract <- merge(extract, redbook2, by.x="ndc", by.y="ndcnum", all.x = TRUE)
names(extract)[names(extract)=="gennme.x"] <- "gennme"
names(extract)[names(extract)=="prodnme.x"] <- "prodnme"

# make sure the individual is excluded if they have used methadone
methadone <- af_extract[which(af_extract$ndc=="H0020"| af_extract$ndc %in% redbook_methadone$ndcnum), ]
# extract <- extract[!(extract$ndc=="H0020" | extract$ndc %in% redbook_methadone$ndcnum), ] 
basic <- basic[!(basic$studyid %in% methadone$StudyID), ] #15272 to 15103 observations


# -132 to -43 days = 90 days
# -42 to -1 days = 42 days

# get days of OUD medications
# 90 days before delivery
# extract90 <- extract[extract$Rx_B90_DOB==1 & extract$Expose1==1, ] # no longer subset on expose1 (per Andrew, January 2026)
extract90 <- extract[extract$Rx_B90_DOB==1, ]
names(extract90) <- tolower(names(extract90))
extract90 <- extract90[extract90$studyid %in% basic$studyid, ]
extract90 <- extract90[order(extract90$studyid, extract90$from), ]

# extract90 <- extract90 %>% group_by(studyid) %>% 
#                   mutate(earliest.rx = min(from),
#                         last.rx = max(from),
#                             next.rx.exp = as.Date(from) + daysup,
#                                 next.rx.actual = lead(from),
#                                   prev.rx = lag(from),
#                                     diff.days.rx = from - prev.rx) 


# if expected next prescription date (date when med will run out is the same in two rows, then delete the second row)
# extract90 <- extract90 %>% group_by(studyid) %>% 
#                   mutate(keep = case_when(prev.rx + lag(daysup) > from + diff.days.rx ~ 0))
# 
# extract90 <- extract90 %>% group_by(studyid) %>% 
#                   mutate(keep = case_when(next.rx.exp == lag(next.rx.exp) ~ 0,
#                                           from == earliest.rx ~ 1, 
#                                           from == last.rx ~ 1))

# extract90$keep <- ifelse(is.na(extract90$keep) & )
# IDs to check 
# extract90[which(extract90$studyid=="002009032034-04544465-03494573"), c("studyid", "daysup", "from", "total.rx.days")]
# "002008060134-05537053-05774374"
# "002008032250-05512799-02853395"

# summarize the number of days supplied cap at 90+42 days = 132 days (Method 1)
extract90 <- extract90 %>% group_by(studyid) %>% 
                  mutate(total.rx.days = cumsum(daysup)) %>% data.frame()

#extract90[which(extract90$studyid=="002009032034-04544465-03494573"), c("studyid", "daysup", "from", "total.rx.days")]
extract90$total.rx.days <- ifelse(extract90$total.rx.days > 90, 90, extract90$total.rx.days)

extract90s <- extract90 %>% group_by(studyid) %>% slice_max(total.rx.days, with_ties = FALSE) %>% data.frame()


# 41 days after delivery
# extract41 <- extract[extract$Rx_DOB_P41==1 & extract$Expose1==1, ] (no longer needed per Andrew, January 2026)
extract41 <- extract[extract$Rx_DOB_P41==1, ]
names(extract41) <- tolower(names(extract41))
extract41 <- extract41[extract41$studyid %in% basic$studyid, ]
extract41 <- extract41[order(extract41$studyid, extract41$from), ]

extract41 <- extract41 %>% group_by(studyid) %>% 
                  mutate(total.rx.days = cumsum(daysup)) %>% data.frame()
extract41$total.rx.days <- ifelse(extract41$total.rx.days > 42, 42, extract41$total.rx.days)

extract41s <- extract41 %>% group_by(studyid) %>% slice_max(total.rx.days, with_ties = FALSE) %>% data.frame()


basic$total.rx.days.b90 <- extract90s[match(basic$studyid, extract90s$studyid), "total.rx.days"]
basic$total.rx.days.p41 <- extract41s[match(basic$studyid, extract41s$studyid), "total.rx.days"]

basic$total.rx.days.b90 <- ifelse(is.na(basic$total.rx.days.b90), 0, basic$total.rx.days.b90)
basic$total.rx.days.b90 <- ifelse(basic$total.rx.days.b90 > 90, 90, basic$total.rx.days.b90)
basic$total.rx.days.p41 <- ifelse(is.na(basic$total.rx.days.p41), 0, basic$total.rx.days.p41)



# Method 2
# Add up total days supplied, if > 90 then add that number to days after delivery instead of capping at 90
extract90_2 <- extract90 %>% group_by(studyid) %>% 
                  mutate(total.rx.days = cumsum(daysup),
                         max.rx.days = max(total.rx.days)) %>% data.frame()

extract90_2$total.rx.minus90 <- ifelse(extract90_2$max.rx.days > 90, extract90_2$max.rx.days - 90, 0)


extract41_2 <- extract41 %>% group_by(studyid) %>% 
                  mutate(total.rx.days = cumsum(daysup),
                         max.rx.days = max(total.rx.days)) %>% data.frame()

basic$total.rx.days.b90_2 <- extract90_2[match(basic$studyid, extract90_2$studyid), "max.rx.days"]
basic$total.rx.days.b90_2 <- ifelse(basic$total.rx.days.b90_2 > 90, 90, basic$total.rx.days.b90_2)
basic$total.rx.days.b90_2 <- ifelse(is.na(basic$total.rx.days.b90_2), 0, basic$total.rx.days.b90_2)
basic$total.rx.days.b90_minus90 <- extract90_2[match(basic$studyid, extract90_2$studyid), "total.rx.minus90"]
basic$total.rx.days.b90_minus90 <- ifelse(is.na(basic$total.rx.days.b90_minus90), 0, basic$total.rx.days.b90_minus90)

basic$total.rx.days.p41_2max <- extract41_2[match(basic$studyid, extract41_2$studyid), "max.rx.days"]
basic$total.rx.days.p41_2max <- ifelse(basic$total.rx.days.p41_2max > 42, 42, basic$total.rx.days.p41_2max)
basic$total.rx.days.p41_2max <- ifelse(is.na(basic$total.rx.days.p41_2max), 0, basic$total.rx.days.p41_2max)

basic$total.rx.days.p41_2 <- rowSums(basic[, c("total.rx.days.p41_2max", "total.rx.days.b90_minus90")])
basic$total.rx.days.p41_2 <- ifelse(basic$total.rx.days.p41_2 > 42, 42, basic$total.rx.days.p41_2)

# Method 3
# counting for overlapping prescription dates
# extract90_3_old <- extract90 %>% group_by(studyid) %>%
#     arrange(studyid, from_day) %>%
#     mutate(indx = c(0, cumsum(as.numeric(lead(from_day)) >
#                      cummax(as.numeric(from_day + daysup)))[-n()])) %>%
#     group_by(studyid, indx) %>%
#     summarise(start = min(from_day), end = max(from_day + daysup))

extract90_3 <- extract90 
extract90_3$strength <- as.numeric(gsub( " .*$", "", extract90_3$strngth))
extract41_3 <- extract41
extract41_3$strength <- as.numeric(gsub( " .*$", "", extract41_3$strngth))

newdf <- bind_rows(extract90_3, extract41_3)
# exclude outlier where qty is 56000
newdf <- newdf[!newdf$qty==56000, ]
# exclude record where days supply is 0
newdf <- newdf[!newdf$daysup==0, ]

# powder buprenorphine ndc: 51927101200; exclude in days supply but not dosage calculation 
# powder bup was taken by someone with Expose1==0, so they are not included 
# exclude sublocade in dosage calculation, but they can be included in the days supply calculation
# same with vivitrol; calculate days supply but not the dosage

# who got injectable (solution) versus not
# who got buprenorphine versus not

newdf_dose <- newdf[!(newdf$prodnme %in% "sublocade 300mg"), ] #only 10 observations differ and 1 study id ("002020059663-06267662-03931614")
newdf_dose <- newdf_dose[newdf_dose$prodnme != "vivitrol", ] #differ by 78 study ids
newdf_dose <- newdf_dose[newdf_dose$prodnme != "sublocade 300mg", ] # this did not reduce anymore people


calc_coverage_and_dose <- function(df, pre_cap = 90, post_cap = 42) {

  # Expand ranges to unique day integers
  get_unique_days <- function(start_vec, end_vec) {
    sort(unique(unlist(purrr::map2(start_vec, end_vec, seq))))
  }

  # Safely coerce possible NULL/NA to integer(0) for set operations
  safe_int <- function(x) {
    if (is.null(x) || (length(x) == 1 && is.na(x))) integer(0) else as.integer(x)
  }

  df <- df %>%
    mutate(
      fill_start = from_day,
      fill_end   = from_day + daysup - 1
    )

  # ---------------- PRE-PERIOD ----------------
  pre_df <- df %>% filter(from_day >= -132, from_day <= -43) %>%
    mutate(
      fill_start = pmax(fill_start, -132),
      fill_end   = fill_start + daysup - 1
    )

  pre_summary <- pre_df %>%
    group_by(studyid) %>%
    summarise(
      all_covered_days = list(get_unique_days(fill_start, fill_end)),
      .groups = "drop"
    ) %>%
    mutate(
      pre_period_days   = map(all_covered_days, ~ .x[.x >= -132 & .x <= -43]),
      capped_days       = map(pre_period_days, ~ head(.x, pre_cap)),
      pre_days_covered  = map_int(capped_days, length),

      # days that didn't fit in capped pre; then restrict to post window for carryover
      carryover_day_list = map2(all_covered_days, capped_days, ~ setdiff(.x, .y)),
      carryover_days_vec = map(carryover_day_list, ~ .x[.x >= -42 & .x <= -1]),
      carryover_days     = map_int(carryover_days_vec, length)
    )

  capped_pre_df <- pre_df %>%
    inner_join(pre_summary, by = "studyid") %>%
    rowwise() %>%
    mutate(
      actual_days        = list(seq(fill_start, fill_end)),
      overlap_days       = length(intersect(actual_days, capped_days)),
      proportion         = overlap_days / daysup,
      adjusted_quantity  = qty * proportion,
      strength_quantity  = qty * strength * proportion
    ) %>%
    ungroup()

  pre_dose_summary <- capped_pre_df %>%
    group_by(studyid) %>%
    summarise(
      total_adjusted_quantity_pre = sum(adjusted_quantity, na.rm = TRUE),
      total_strength_quantity_pre = sum(strength_quantity, na.rm = TRUE),
      total_daysup_pre            = sum(daysup * proportion, na.rm = TRUE),
      avg_dose_pre = if_else(total_daysup_pre > 0,
                             total_strength_quantity_pre / total_daysup_pre, 0),
      .groups = "drop"
    )

  pre_summary <- pre_summary %>%
    left_join(pre_dose_summary, by = "studyid")

  # ---------------- POST-PERIOD ----------------
  post_df <- df %>% filter(from_day >= -42, from_day <= -1) %>%
    mutate(
      fill_start = pmax(fill_start, -42),
      fill_end   = fill_start + daysup - 1
    )

  post_summary <- post_df %>%
    group_by(studyid) %>%
    summarise(
      all_covered_days = list(get_unique_days(fill_start, fill_end)),
      .groups = "drop"
    ) %>%
    mutate(
      post_period_days  = map(all_covered_days, ~ .x[.x >= -42 & .x <= -1]),
      post_capped_days  = map(post_period_days, ~ head(.x, post_cap)),
      post_days_covered = map_int(post_capped_days, length),
      effective_days    = post_days_covered
    )

  capped_post_df <- post_df %>%
    inner_join(post_summary, by = "studyid") %>%
    rowwise() %>%
    mutate(
      actual_days        = list(seq(fill_start, fill_end)),
      overlap_days       = length(intersect(actual_days, post_capped_days)),
      proportion         = overlap_days / daysup,
      adjusted_quantity  = qty * proportion,
      strength_quantity  = qty * strength * proportion
    ) %>%
    ungroup()

  post_dose_summary <- capped_post_df %>%
    group_by(studyid) %>%
    summarise(
      total_adjusted_quantity_post = sum(adjusted_quantity, na.rm = TRUE),
      total_strength_quantity_post = sum(strength_quantity, na.rm = TRUE),
      total_daysup_post            = sum(daysup * proportion, na.rm = TRUE),
      avg_dose_post = if_else(total_daysup_post > 0,
                              total_strength_quantity_post / total_daysup_post, 0),
      .groups = "drop"
    )

  post_summary <- post_summary %>%
    left_join(post_dose_summary, by = "studyid")

  # ---------------- COMBINE PRE + POST ----------------
  full_summary <- pre_summary %>%
    select(studyid, capped_days, carryover_days_vec, pre_days_covered, carryover_days,
           total_adjusted_quantity_pre, total_strength_quantity_pre,
           total_daysup_pre, avg_dose_pre) %>%
    full_join(
      post_summary %>%
        select(studyid, post_capped_days, post_days_covered, effective_days,
               total_adjusted_quantity_post, total_strength_quantity_post,
               total_daysup_post, avg_dose_post),
      by = "studyid"
    ) %>%
    mutate(across(
      c(pre_days_covered, carryover_days, total_adjusted_quantity_pre,
        total_strength_quantity_pre, total_daysup_pre, avg_dose_pre,
        post_days_covered, effective_days, total_adjusted_quantity_post,
        total_strength_quantity_post, total_daysup_post, avg_dose_post),
      ~ replace_na(.x, 0)
    )) %>%
    mutate(
      total_days_covered           = pre_days_covered + post_days_covered,
      total_adjusted_quantity_all  = total_adjusted_quantity_pre + total_adjusted_quantity_post,
      total_strength_quantity_all  = total_strength_quantity_pre + total_strength_quantity_post,
      total_daysup_all             = total_daysup_pre + total_daysup_post,
      avg_dose_entire_period       = if_else(total_daysup_all > 0,
                                             total_strength_quantity_all / total_daysup_all, 0),

      # NEW: unique post-period days including carryover spillover
      total_unique_post_days_incl_carryover = map2_int(
        carryover_days_vec, post_capped_days,
        ~ length(unique(c(safe_int(.x), safe_int(.y))))
      ),

      # NEW: true total unique days supplied across periods (pre capped + carryover + post capped)
      total_unique_days_supplied = pmap_int(
        list(capped_days, carryover_days_vec, post_capped_days),
        function(a, b, c) length(unique(c(safe_int(a), safe_int(b), safe_int(c))))
      )
    ) %>%
    # drop helper list-columns from final output (keep if you want to audit)
    select(-capped_days, -carryover_days_vec, -post_capped_days)

  return(full_summary)
}

# manuscript example: id 002009013002-04514929-02081695 (old); now use: (002007035729-02903621-04746761)
# total_daysup_pre adds the overlapping/duplicate rx dates
# use pre_days_covered and post_days_covered (adjusted)
# need to take into account for carryover_days that do not overlap in the post-period and add that to post_days_covered

extract_method3 <- calc_coverage_and_dose(newdf)
extract_method3 <- as.data.frame(extract_method3)
extract <- extract[order(extract$StudyID, extract$from_day),]

# to describe dose use newdf_dose
extract_method3_dose <- calc_coverage_and_dose(newdf_dose)
extract_method3_dose <- as.data.frame(extract_method3_dose)


# check

# extract[which(extract$StudyID=="002012002626-01447011-05711550"), c("StudyID", "from_day", "daysup", "qty", "strngth")]
# extract[which(extract$StudyID=="002012081578-01529551-05299669"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]
# extract[which(extract$StudyID=="002007047985-02911169-05614828"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]
# extract[which(extract$StudyID=="002007035729-02903621-04746761"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]
# extract[which(extract$StudyID=="002007050401-02350074-04750282"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]
# extract[which(extract$StudyID=="002009013002-04514929-02081695"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]

# merge total days and dose
extract_method3 <- as.data.frame(extract_method3)
# pre-days covered
basic$total.rx.days.b90_3 <- extract_method3[match(basic$studyid, extract_method3$studyid), "pre_days_covered"]
# post-days covered
basic$total.rx.days.p41_3 <- extract_method3[match(basic$studyid, extract_method3$studyid), "total_unique_post_days_incl_carryover"]
# total days covered
# basic$total.rx.days.total_3 <- extract_method3[match(basic$studyid, extract_method3$studyid), "total_unique_days_supplied"]

#total days shifted
calc_coverage_and_dose_shifted <- function(df, pre_cap = 90, post_cap = 42) {
  
  library(dplyr)
  library(purrr)
  library(tidyr)
  
  # ---------------- HELPERS ----------------
  get_unique_days <- function(start_vec, end_vec) {
    sort(unique(unlist(map2(start_vec, end_vec, seq))))
  }
  
  safe_int <- function(x) {
    if (is.null(x) || (length(x) == 1 && is.na(x))) integer(0) else as.integer(x)
  }
  
  # NEW: shifted-days helper (sequential, no overlap)
  calc_shifted_days <- function(from_day, daysup,
                                window_start = -132,
                                window_end   = -1) {
    
    ord <- order(from_day)
    from_day <- from_day[ord]
    daysup   <- daysup[ord]
    
    current_day <- -Inf
    total_days  <- 0
    
    for (i in seq_along(from_day)) {
      start_day <- max(from_day[i], current_day)
      end_day   <- start_day + daysup[i] - 1
      
      usable_end <- min(end_day, window_end)
      usable_days <- max(0, usable_end - start_day + 1)
      
      total_days <- total_days + usable_days
      current_day <- end_day + 1
    }
    
    total_days
  }
  
  # ---------------- BASE DF ----------------
  df <- df %>%
    mutate(
      fill_start = from_day,
      fill_end   = from_day + daysup - 1
    )
  
  # ---------------- PRE-PERIOD ----------------
  pre_df <- df %>%
    filter(from_day >= -132, from_day <= -43) %>%
    mutate(
      fill_start = pmax(fill_start, -132),
      fill_end   = fill_start + daysup - 1
    )
  
  pre_summary <- pre_df %>%
    group_by(studyid) %>%
    summarise(
      all_covered_days = list(get_unique_days(fill_start, fill_end)),
      .groups = "drop"
    ) %>%
    mutate(
      pre_period_days   = map(all_covered_days, ~ .x[.x >= -132 & .x <= -43]),
      capped_days       = map(pre_period_days, ~ head(.x, pre_cap)),
      pre_days_covered  = map_int(capped_days, length),
      
      carryover_day_list = map2(all_covered_days, capped_days, ~ setdiff(.x, .y)),
      carryover_days_vec = map(carryover_day_list, ~ .x[.x >= -42 & .x <= -1]),
      carryover_days     = map_int(carryover_days_vec, length)
    )
  
  capped_pre_df <- pre_df %>%
    inner_join(pre_summary, by = "studyid") %>%
    rowwise() %>%
    mutate(
      actual_days       = list(seq(fill_start, fill_end)),
      overlap_days      = length(intersect(actual_days, capped_days)),
      proportion        = overlap_days / daysup,
      adjusted_quantity = qty * proportion,
      strength_quantity = qty * strength * proportion
    ) %>%
    ungroup()
  
  pre_dose_summary <- capped_pre_df %>%
    group_by(studyid) %>%
    summarise(
      total_adjusted_quantity_pre = sum(adjusted_quantity, na.rm = TRUE),
      total_strength_quantity_pre = sum(strength_quantity, na.rm = TRUE),
      total_daysup_pre            = sum(daysup * proportion, na.rm = TRUE),
      avg_dose_pre = if_else(total_daysup_pre > 0,
                             total_strength_quantity_pre / total_daysup_pre, 0),
      .groups = "drop"
    )
  
  pre_summary <- pre_summary %>%
    left_join(pre_dose_summary, by = "studyid")
  
  # ---------------- POST-PERIOD ----------------
  post_df <- df %>%
    filter(from_day >= -42, from_day <= -1) %>%
    mutate(
      fill_start = pmax(fill_start, -42),
      fill_end   = fill_start + daysup - 1
    )
  
  post_summary <- post_df %>%
    group_by(studyid) %>%
    summarise(
      all_covered_days = list(get_unique_days(fill_start, fill_end)),
      .groups = "drop"
    ) %>%
    mutate(
      post_period_days  = map(all_covered_days, ~ .x[.x >= -42 & .x <= -1]),
      post_capped_days  = map(post_period_days, ~ head(.x, post_cap)),
      post_days_covered = map_int(post_capped_days, length),
      effective_days    = post_days_covered
    )
  
  capped_post_df <- post_df %>%
    inner_join(post_summary, by = "studyid") %>%
    rowwise() %>%
    mutate(
      actual_days       = list(seq(fill_start, fill_end)),
      overlap_days      = length(intersect(actual_days, post_capped_days)),
      proportion        = overlap_days / daysup,
      adjusted_quantity = qty * proportion,
      strength_quantity = qty * strength * proportion
    ) %>%
    ungroup()
  
  post_dose_summary <- capped_post_df %>%
    group_by(studyid) %>%
    summarise(
      total_adjusted_quantity_post = sum(adjusted_quantity, na.rm = TRUE),
      total_strength_quantity_post = sum(strength_quantity, na.rm = TRUE),
      total_daysup_post            = sum(daysup * proportion, na.rm = TRUE),
      avg_dose_post = if_else(total_daysup_post > 0,
                              total_strength_quantity_post / total_daysup_post, 0),
      .groups = "drop"
    )
  
  post_summary <- post_summary %>%
    left_join(post_dose_summary, by = "studyid")
  
  # ---------------- SHIFTED DAYS (NEW) ----------------
  shifted_days <- df %>%
    group_by(studyid) %>%
    summarise(
      total_shifted_days_supplied =
        calc_shifted_days(from_day, daysup),
      .groups = "drop"
    )
  
  # ---------------- COMBINE ----------------
  full_summary <- pre_summary %>%
    select(studyid, capped_days, carryover_days_vec,
           pre_days_covered, carryover_days,
           total_adjusted_quantity_pre,
           total_strength_quantity_pre,
           total_daysup_pre, avg_dose_pre) %>%
    full_join(
      post_summary %>%
        select(studyid, post_capped_days,
               post_days_covered, effective_days,
               total_adjusted_quantity_post,
               total_strength_quantity_post,
               total_daysup_post, avg_dose_post),
      by = "studyid"
    ) %>%
    left_join(shifted_days, by = "studyid") %>%
    mutate(across(
      c(pre_days_covered, carryover_days,
        total_adjusted_quantity_pre,
        total_strength_quantity_pre,
        total_daysup_pre, avg_dose_pre,
        post_days_covered, effective_days,
        total_adjusted_quantity_post,
        total_strength_quantity_post,
        total_daysup_post, avg_dose_post,
        total_shifted_days_supplied),
      ~ replace_na(.x, 0)
    )) %>%
    mutate(
      total_days_covered          = pre_days_covered + post_days_covered,
      total_adjusted_quantity_all =
        total_adjusted_quantity_pre + total_adjusted_quantity_post,
      total_strength_quantity_all =
        total_strength_quantity_pre + total_strength_quantity_post,
      total_daysup_all =
        total_daysup_pre + total_daysup_post,
      avg_dose_entire_period =
        if_else(total_daysup_all > 0,
                total_strength_quantity_all / total_daysup_all, 0),
      
      total_unique_days_supplied = pmap_int(
        list(capped_days, carryover_days_vec, post_capped_days),
        function(a, b, c)
          length(unique(c(safe_int(a), safe_int(b), safe_int(c))))
      )
    ) %>%
    select(-capped_days, -carryover_days_vec, -post_capped_days)
  
  return(full_summary)
}

extract_method3_shifted <- calc_coverage_and_dose_shifted(newdf)


extract_method3_shifted <- as.data.frame(extract_method3_shifted)
basic$total.rx.days.shifted <- extract_method3_shifted[match(basic$studyid, extract_method3_shifted$studyid), "total_shifted_days_supplied"]
label(basic$total.rx.days.shifted) <- "MOUD days exposure shifted approach"

# labels
label(basic$total.rx.days.b90) <- "Total days supplied MOUD before delivery"
label(basic$total.rx.days.p41) <- "Total days supplied MOUD after delivery"

label(basic$total.rx.days.b90_2) <- "Total days supplied MOUD before delivery (method 2)"
label(basic$total.rx.days.p41_2) <- "Total days supplied MOUD after delivery (method 2)"

label(basic$total.rx.days.b90_3) <- "Total days supplied MOUD before delivery (method 3)"
label(basic$total.rx.days.p41_3) <- "Total days supplied MOUD after delivery (method 3)"


basic$avg.dose.pre <- extract_method3_dose[match(basic$studyid, extract_method3_dose$studyid), "avg_dose_pre"]
basic$avg.dose.post <- extract_method3_dose[match(basic$studyid, extract_method3_dose$studyid), "avg_dose_post"]
basic$avg.dose.total <- extract_method3_dose[match(basic$studyid, extract_method3_dose$studyid), "avg_dose_entire_period"]

label(basic$avg.dose.pre) <- "Avg dose pre-delivery (mg)"
label(basic$avg.dose.post) <- "Avg dose post-delivery (mg)"
label(basic$avg.dose.total) <- "Avg dose exposure (mg)"

# # summarize very last medication received per person
# extract_last <- extract %>% group_by(StudyID) %>% slice_max(from_day) %>% data.frame
# names(extract_last) <- tolower(names(extract_last))
# ## summarize the type of medication and number of prescriptions 
# extract_last <- extract_last[ , c("studyid", "from_day", "gennme")]
# extract_last <- unique(extract_last)
# extract_last <- reshape2::dcast(extract_last, studyid ~ gennme)
# extract_last$last_rx_type <- apply(extract_last[, 2:7], 1, function(x) toString(na.omit(x)))
# names(extract_last) <- make.names(names(extract_last))

## summarize the type of medication and number of prescriptions 
extract_u <- extract
names(extract_u) <- tolower(names(extract_u))
extract_u <- extract_u[ , c("studyid", "from_day", "gennme")]
extract_u <- unique(extract_u)
extract_u <- reshape2::dcast(extract_u, studyid ~ gennme)
names(extract_u) <- make.names(names(extract_u))
extract_u <- extract_u %>%  mutate_if(is.numeric, ~1 * (. > 0))
extract_u$unique.rx <- rowSums(extract_u[ , 2:6], na.rm = TRUE)
# filter out people in our dataset only 
extract_u <- extract_u[which(extract_u$studyid %in% basic$studyid),]
# excluded outlier with qty of 56000 so exclude this person 
extract_u <- extract_u[which(extract_u$studyid != "002020030193-06235149-04443003"), ]
# summarize how many took which medications
extract_u$bup <- ifelse(extract_u$Buprenorphine==1 | extract_u$Buprenorphine.Hydrochloride==1 | extract_u$Buprenorphine.Naloxone==1, 1, 0)
extract_u$nal <- ifelse(extract_u$Naltrexone==1 | extract_u$Naltrexone.Hydrochloride==1, 1, 0)
extract_u$rxtype <- ifelse(extract_u$bup==1 & extract_u$nal==0, "Buprenorphine only", 
                           ifelse(extract_u$nal==1 & extract_u$bup==0, "Naltrexone only",
                                                                          ifelse(extract_u$bup==1 & extract_u$nal==1, "Both", NA)))


# basic$last_rx_type <- extract_last[match(basic$studyid, extract_last$studyid), "last_rx_type"]
basic$unique.rx <- extract_u[match(basic$studyid, extract_u$studyid), "unique.rx"]
basic$rxtype <- extract_u[match(basic$studyid, extract_u$studyid), "rxtype"]
label(basic$rxtype) <- "Type of MOUD"

# 
# basic$last_rx_type <- as.factor(basic$last_rx_type)
# label(basic$last_rx_type) <- "Last medication type"

label(basic$unique.rx) <- "Unique number of prescriptions"


# dosage and exposure days information without naltrexone and methadone
newdf2 <- newdf
newdf2 <- newdf2[!(newdf2$gennme %in% c("Naltrexone", "Naltrexone Hydrochloride")), ]

newdf_dose2 <- newdf_dose
newdf_dose2 <- newdf_dose2[!(newdf_dose2$gennme %in% c("Naltrexone", "Naltrexone Hydrochloride")), ]

extract_method3_bup <- calc_coverage_and_dose(newdf2)
extract_method3_bup <- as.data.frame(extract_method3_bup)

extract_method3_bup_dose <- calc_coverage_and_dose(newdf_dose2)
extract_method3_bup_dose <- calc_coverage_and_dose(newdf_dose2)

# check

# extract[which(extract$StudyID=="002012002626-01447011-05711550"), c("StudyID", "from_day", "daysup", "qty", "strngth")]
# extract[which(extract$StudyID=="002012081578-01529551-05299669"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]
# extract[which(extract$StudyID=="002007047985-02911169-05614828"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]
# extract[which(extract$StudyID=="002018049083-01868012-03765766"), c("StudyID", "from_day", "daysup",  "qty", "strngth")]

# merge total days and dose
extract_method3_bup <- as.data.frame(extract_method3_bup)
# pre-days covered
basic$total.rx.days.b90_3bup <- extract_method3_bup[match(basic$studyid, extract_method3_bup$studyid), "pre_days_covered"]
# post-days covered
basic$total.rx.days.p41_3bup <- extract_method3_bup[match(basic$studyid, extract_method3_bup$studyid), "total_unique_post_days_incl_carryover"]
# total days covered
basic$total.rx.days.total_3bup <- extract_method3_bup[match(basic$studyid, extract_method3_bup$studyid), "total_unique_days_supplied"]

label(basic$total.rx.days.b90_3bup) <- "Total days buprenorphine before delivery (method 3)"
label(basic$total.rx.days.p41_3bup) <- "Total days buprenorphine after delivery (method 3)"

extract_method3_bup_dose <- as.data.frame(extract_method3_bup_dose)
basic$avg.dose.pre.bup <- extract_method3_bup_dose[match(basic$studyid, extract_method3_bup_dose$studyid), "avg_dose_pre"]
basic$avg.dose.post.bup <- extract_method3_bup_dose[match(basic$studyid, extract_method3_bup_dose$studyid), "avg_dose_post"]
basic$avg.dose.total.bup <- extract_method3_bup_dose[match(basic$studyid, extract_method3_bup_dose$studyid), "avg_dose_entire_period"]

label(basic$avg.dose.pre.bup) <- "Avg dose BUP pre-delivery (mg)"
label(basic$avg.dose.post.bup) <- "Avg dose BUP post-delivery (mg)"
label(basic$avg.dose.total.bup) <- "Avg dose BUP exposure (mg)"

```

```{r, fig.height=3, fig.width=11}
# ID: 002009013002-04514929-02081695 (old)
# ID: 002007035729-02903621-04746761 (new)
lines_data <- data.frame(
  line_id = factor(1:4),
  x_start = c(-50, -46, -43, -40),
  x_end   = c(-23, -44, -42, -14),
  y       = c(4, 3, 2, 1)  # vertical positions for each line
)

# Plot
custom_labels <- c("Rx 4", "Rx 3", "Rx 2", "Rx 1")
bottom_labels <- c("28 unique days", "0 unique days", "0 unique days", "9 unique days") 

p1 <- ggplot(lines_data) +
  geom_segment(aes(x = x_start, xend = x_end, y = y, yend = y), 
               size = 2, color = "black") +
  # Annotate start values
  geom_text(aes(x = x_start, y = y + 0.1, label = x_start), 
            size = 4, color = "black", hjust = 1) +
  # Annotate end values
  geom_text(aes(x = x_end, y = y + 0.1, label = x_end), 
            size = 4, color = "black", hjust = 0) +
  # Annotate custom values below the line
  geom_text(aes(x = (x_start + x_end)/2, y = y - 0.2, label = bottom_labels), size = 4, color = "black") +
  geom_vline(xintercept = -1, linetype = "dotted", color = "red", size = 1) +
  scale_x_continuous(breaks = seq(-50, 7, 10)) +
  scale_y_continuous(breaks = 1:4, labels= custom_labels) +
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "Non-overlapping days MOUD", x = "Day", y = "")

# same ID but using "shift-forward" approach

lines_data2 <- data.frame(
  line_id = factor(1:4),
  x_start = c(-50, -22, -19, -17),
  x_end   = c(-23, -20, -18, 9),
  y       = c(4, 3, 2, 1)  # vertical positions for each line
)

# Plot
custom_labels2 <- c("Rx 4", "Rx 3", "Rx 2", "Rx 1")
bottom_labels2 <- c("28 unique days", "3 unique days", "2 unique days", "17 unique days") 

p2 <- ggplot(lines_data2) +
  geom_segment(aes(x = x_start, xend = x_end, y = y, yend = y), 
               size = 2, color = "black") +
  # Annotate start values
  geom_text(aes(x = x_start, y = y + 0.1, label = x_start), 
            size = 4, color = "black", hjust = 1) +
  # Annotate end values
  geom_text(aes(x = x_end, y = y + 0.1, label = x_end), 
            size = 4, color = "black", hjust = 0) +
  # Annotate custom values below the line
  geom_text(aes(x = (x_start + x_end)/2, y = y - 0.2, label = bottom_labels2), size = 4, color = "black") +
  geom_vline(xintercept = -1, linetype = "dotted", color = "red", size = 1) +
  scale_x_continuous(breaks = seq(-50, 7, 10)) +
  scale_y_continuous(breaks = 1:4, labels= custom_labels2) +
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "Shift-forward days MOUD", x = "Day", y = "")


ggpubr::ggarrange(p1, p2,
                  ncol=2, nrow = 1)

```



```{r}
# follow-up time if they had an event (counting events not captured in claims using all hospital payer data)
# if death or overdose happens after 365 days, then use overdose day or last observed day
af_out365 <- af_out[(af_out$momdeath_day >= 0 & af_out$momdeath_day <= 323) | (af_out$overdose_opiod_day >=0 & af_out$overdose_opiod_day <= 323 ) |  (af_out$fdate_seg1_day >=0 & af_out$fdate_seg1_day <= 323) | (af_out$overdose_nonopiod_day >=0 & af_out$overdose_nonopiod_day <= 323), ]

# exclude if they had OD in both exposure and outcome
af_out365 <- af_out365[!(af_out365$overdose_nonopioid_exposure==1 & af_out365$overdose_nonopioid_outcome==1), ]
af_out365 <- af_out365[!(af_out365$overdose_opioid_exposure==1 & af_out365$overdose_opioid_outcome==1), ]
af_out365 <- af_out365[!(af_out365$overdose_nonopioid_exposure==1 & af_out365$overdose_opioid_outcome==1), ]
af_out365 <- af_out365[!(af_out365$overdose_opioid_exposure==1 & af_out365$overdose_nonopioid_outcome==1), ]


af_out365$overdose_day <- pmin(af_out365$overdose_nonopiod_day, af_out365$overdose_opiod_day, na.rm = TRUE)
# af_out365 <- af_out365 %>% group_by(studyid) %>% slice_min(overdose_day, with_ties = FALSE) %>% data.frame()

af_out365$days.to.event <- pmin(af_out365$momdeath_day, af_out365$overdose_day, na.rm = TRUE)
af_out365$outcome <- ifelse(af_out365$days.to.event== af_out365$momdeath_day , "Death", 
                            ifelse(af_out365$days.to.event == af_out365$overdose_day, "Overdose", NA))
af_out365$outcome <- ifelse(is.na(af_out365$outcome) & is.na(af_out365$momdeath_day) & !is.na(af_out365$overdose_day), "Overdose", af_out365$outcome)

af_out365 <- as.data.frame(af_out365)
af_out <- as.data.frame(af_out)
af_out365$unique_id <- sub(".*-", "", af_out365$studyid)
af_out$unique_id <- sub(".*-", "", af_out$studyid)

# match on study ID to count overdoses that occurred in subsequent pregnancies
af_out$days.to.event <- af_out365[match(af_out$studyid, af_out365$studyid), "days.to.event"]
basic$days.to.event <- af_out365[match(basic$studyid, af_out365$studyid), "days.to.event"]
basic$outcome <- af_out365[match(basic$studyid, af_out365$studyid), "outcome"]

# af_out365[which(!is.na(af_out365$momdeath_day) & !is.na(af_out365$overdose_day)), c("studyid", "overdose_day", "fdate_seg1_day", "momdeath_day", "days.to.event", "outcome")]

# basic$days.to.event <- ifelse(is.na(basic$days.to.event), basic$fdate_seg1_day, basic$days.to.event)
# change so that if loss of enrollment occurs, we assume that no event happened if they are not found in death certificate or HDDS (change follow-up to 323 days or 1 year postpartum)
basic$days.to.event <- ifelse(is.na(basic$days.to.event), 323, basic$days.to.event)
basic$days.to.event <- ifelse(basic$days.to.event > 323, 323, basic$days.to.event)

# to check
# basic[which(basic$fdate_seg1_day < basic$days.to.event), c("studyid", "death.or.od",  "days.to.event", "fdate_seg1_day")]
# these people had fdate_seg1_day before death or overdose date; ???

# All dates are relative to Day 42, so everything is shifted to day-42
basic$death.day.from.deliv = basic$death_day + 42
basic$days.to.event.from.deliv = basic$days.to.event + 42
basic$fdate.from.deliv = basic$fdate_seg1_day + 42

# only want outcomes up to 1 year post delivery, which is day 323 from 42 days of delivery
basic <- basic[basic$days.to.event.from.deliv <= 365, ]


# Loss of enrollment within 1 year of delivery
basic$loss.enroll <- ifelse(basic$cat1_seg1==1 & basic$fdate.from.deliv <=365, 1, 0)
label(basic$loss.enroll) <- "Loss of enrollment within 1-year of delivery"

basic$outcome.num <- ifelse(is.na(basic$outcome), 0, 1)


basic$final.outcome <- ifelse(basic$outcome =="Death", 1,
                              ifelse(basic$outcome=="Overdose", 2, 
                                     ifelse(basic$loss.enroll==1 & is.na(basic$outcome), 3, NA)))
basic$final.outcome <- ifelse(is.na(basic$final.outcome) & basic$loss.enroll==1, 3, basic$final.outcome)
basic$final.outcome <- ifelse(is.na(basic$final.outcome) & basic$loss.enroll!=1, 4, basic$final.outcome)


basic$final.outcome.f <- factor(basic$final.outcome, levels=1:4, labels=c("Death", "Overdose", "Loss of enrollment", "Censored"))
label(basic$final.outcome.f) <- "Final outcome"


```


```{r}
# follow-up times calculated without accounting for OD after disenrollment 
af_out365$days.to.event2 <- pmin(af_out365$overdose_day, af_out365$fdate_seg1_day, af_out365$momdeath_day)
af_out365$days.to.event2 <- ifelse(is.na(af_out365$overdose_day), pmin(af_out365$fdate_seg1_day, af_out365$momdeath_day), af_out365$days.to.event2)
af_out365$days.to.event2 <- ifelse(is.na(af_out365$momdeath_day), pmin(af_out365$fdate_seg1_day, af_out365$overdose_day), af_out365$days.to.event2)

af_out365$outcome2 <- ifelse(af_out365$days.to.event2== af_out365$momdeath_day , "Death", 
                            ifelse(af_out365$days.to.event2 == af_out365$overdose_day, "Overdose", NA))
af_out365$outcome2 <- ifelse(is.na(af_out365$outcome2) & is.na(af_out365$momdeath_day) & !is.na(af_out365$overdose_day), "Overdose", af_out365$outcome2)
af_out365$outcome2 <- ifelse(af_out365$days.to.event2 == af_out365$fdate_seg1_day, "Loss of enrollment", af_out365$outcome2)
# check 
# af_out365[which(af_out365$days.to.event != af_out365$days.to.event2), ]

basic$outcome2 <- af_out365[match(basic$studyid, af_out365$studyid), "outcome2"]
basic$days.to.event2 <- af_out365[match(basic$studyid, af_out365$studyid), "days.to.event2"]
basic$days.to.event2 <- ifelse(is.na(basic$days.to.event2), 365, basic$days.to.event2) # if missing then no event; change to max follow-up
basic$days.to.event2 <- ifelse(basic$days.to.event2 > 323, 323, basic$days.to.event) #follow up anchored to day 41 postpartum so event should be 323 days post partum

basic$final.outcome2 <- ifelse(basic$outcome2 =="Death", 1,
                              ifelse(basic$outcome2=="Overdose", 2, 
                                     ifelse(basic$loss.enroll==1 & is.na(basic$outcome2), 3, NA)))
basic$final.outcome2 <- ifelse(is.na(basic$final.outcome2) & basic$loss.enroll==1, 3, basic$final.outcome2)
basic$final.outcome2 <- ifelse(is.na(basic$final.outcome2) & basic$loss.enroll!=1, 4, basic$final.outcome2)

basic$final.outcome2.f <- factor(basic$final.outcome2, levels=1:4, labels=c("Death", "Overdose", "Loss of enrollment", "Censored"))
label(basic$final.outcome2.f) <- "Final outcome (w/o all payer data)"

# to check
# af_out365[which(!is.na(af_out365$momdeath_day) & !is.na(af_out365$overdose_day)), c("studyid", "overdose_day", "fdate_seg1_day", "momdeath_day", "days.to.event", "outcome", "outcome2", "days.to.event2")]
```



```{r}
# Additional covariates to be adjusted for
basic$delivery_year <- as.integer(basic$delivery_year)
label(basic$delivery_year) <- "Year of delivery"

basic$delivery.year.cat <- as.factor(basic$delivery_year)
label(basic$delivery.year.cat) <- "Year of delivery as categorical"

basic$momraceeth_comb.new <- as.character(basic$momraceeth.new)
basic$momraceeth_comb.new <- ifelse(basic$momraceeth_comb.new %nin% c("Non-Hispanic White"), "All Other", basic$momraceeth_comb.new)
basic$momraceeth_comb.new <- as.factor(basic$momraceeth_comb.new)
label(basic$momraceeth_comb.new) <- "Race/Ethnicity binary"

label(basic$total.rx.days.b90_2) <- "Total days supplied MOUD before delivery (method 2)"
label(basic$total.rx.days.p41_2) <- "Total days supplied MOUD after delivery (method 2)"

label(basic$total.rx.days.b90_3) <- "Total days supplied MOUD before delivery (method 3)"
label(basic$total.rx.days.p41_3) <- "Total days supplied MOUD after delivery (method 3)"

basic$total.rx.days.exp <- rowSums(basic[ , c("total.rx.days.b90", "total.rx.days.p41")]) 
basic$total.rx.days.exp2 <- rowSums(basic[ , c("total.rx.days.b90_2", "total.rx.days.p41_2")]) 
basic$total.rx.days.exp3 <- extract_method3[match(basic$studyid, extract_method3$studyid), "total_unique_days_supplied"]

basic$moud.pre <- ifelse(basic$total.rx.days.b90 >0, 1, 0)
basic$moud.pre2 <- ifelse(basic$total.rx.days.b90_2 >0, 1, 0)
basic$moud.pre3 <- ifelse(basic$total.rx.days.b90_3 >0, 1, 0)

basic$moud.post <- ifelse(basic$total.rx.days.p41 >0, 1, 0)
basic$moud.post2 <- ifelse(basic$total.rx.days.p41_2 >0, 1, 0)
basic$moud.post3 <- ifelse(basic$total.rx.days.p41_3 >0, 1, 0)

basic$moud.both <- ifelse(basic$moud.pre==1 & basic$moud.post==1, 1, NA)
basic$moud.both <- ifelse(is.na(basic$moud.both) & (basic$moud.pre==1 | basic$moud.post==1), 2, basic$moud.both)
basic$moud.both <- ifelse(is.na(basic$moud.both), 0, basic$moud.both)

basic$moud.both2 <- ifelse(basic$moud.pre2==1 & basic$moud.post2==1, 1, NA)
basic$moud.both2 <- ifelse(is.na(basic$moud.both2) & (basic$moud.pre2==1 | basic$moud.post2==1), 2, basic$moud.both2)
basic$moud.both2 <- ifelse(is.na(basic$moud.both2), 0, basic$moud.both2)

basic$moud.both3 <- ifelse(basic$moud.pre3==1 & basic$moud.post3==1, 1, NA)
basic$moud.both3 <- ifelse(is.na(basic$moud.both3) & (basic$moud.pre3==1 | basic$moud.post3==1), 2, basic$moud.both3)
basic$moud.both3 <- ifelse(is.na(basic$moud.both3), 0, basic$moud.both3)

basic$moud.onlypre3 <- ifelse(basic$moud.pre3==1 & basic$moud.post3==0, 1, 0)
basic$moud.onlypost3 <- ifelse(basic$moud.post3==1 & basic$moud.pre3==0, 1, 0)

basic$moud.both3div <- ifelse(basic$moud.onlypre3==1, 1, 
                              ifelse(basic$moud.onlypost3==1, 2, 
                                     ifelse(basic$moud.both3==1, 3, 3)))
basic$moud.both3div <- ifelse(is.na(basic$moud.both3div), 0, basic$moud.both3div)
basic$moud.yes <- ifelse(basic$moud.both3div != 0, 1, 0)
basic$moud.yes.f <- factor(basic$moud.yes, levels=0:1, labels=c("No MOUD", "MOUD"))
label(basic$moud.yes.f) <- "MOUD Receipt"
basic$moud.both3div <- factor(basic$moud.both3div, levels=0:3, labels=c("No MOUD", "Only pre", "Only post", "Both pre and post"))

label(basic$moud.both3div) <- "MOUD timing (method 3)"

basic$moud.pre <- factor(basic$moud.pre, levels=0:1, labels=c("No MOUD", "MOUD"))
label(basic$moud.pre) <- "MOUD pre-delivery (method 1)"

basic$moud.pre2 <- factor(basic$moud.pre2, levels=0:1, labels=c("No MOUD", "MOUD"))
label(basic$moud.pre2) <- "MOUD pre-delivery (method 2)"

basic$moud.pre3 <- factor(basic$moud.pre3, levels=0:1, labels=c("No MOUD", "MOUD"))
label(basic$moud.pre3) <- "MOUD pre-delivery (method 3)"

basic$moud.post <- factor(basic$moud.post, levels=0:1, labels=c("No MOUD", "MOUD"))
label(basic$moud.post) <- "MOUD post-delivery (method 1)"

basic$moud.post2 <- factor(basic$moud.post2, levels=0:1, labels=c("No MOUD", "MOUD"))
label(basic$moud.post2) <- "MOUD post-delivery (method 2)"

basic$moud.post3 <- factor(basic$moud.post3, levels=0:1, labels=c("No MOUD", "MOUD"))
label(basic$moud.post3) <- "MOUD post-delivery (method 3)"

# 
basic$moud.both <- factor(basic$moud.both, levels=0:2, labels=c("No MOUD", "Both pre and post", "Either pre or post"))
label(basic$moud.both) <- "MOUD pre and post (method 1)"

basic$moud.both2 <- factor(basic$moud.both2, levels=0:2, labels=c("No MOUD", "Both pre and post", "Either pre or post"))
label(basic$moud.both2) <- "MOUD pre and post (method 2)"

basic$moud.both3 <- factor(basic$moud.both3, levels=0:2, labels=c("No MOUD", "Both pre and post", "Either pre or post"))
label(basic$moud.both3) <- "MOUD pre and post (method 3)"


label(basic$total.rx.days.exp2) <- "Total days supplied MOUD (method 2)"
label(basic$total.rx.days.exp3) <- "Total days supplied MOUD (method 3)"
# 

basic$anx.dep.comb <- ifelse(basic$anxiety==1 | basic$depression== 1, 1, 0)
basic$anx.dep.comb <- ifelse(is.na(basic$anx.dep.comb), 0, basic$anx.dep.comb)
basic$anx.dep.comb.f <- factor(basic$anx.dep.comb, levels=0:1, labels=c("No", "Yes"))
label(basic$anx.dep.comb.f) <- "Anxiety and depression combined"

basic$outcome.bin <- ifelse(basic$final.outcome %in% c(1, 2), 1, 0)

```



# Introduction 

<br>

This project focuses on Analysis 2 of the Study Protocol, which aims to characterize pregnant individuals with OUD and examine hazards of maternal death and overdose. A poster was presented at Academy Health annual research meeting focusing on analysis on those with OUD and at least one day of MOUD prescription. (Section 7)

## Methods

<br>

Data Source

+ TennCare data
+ TN all payer hospitalization database
+ Birth and death certificates
+ US Census data

<br>

Inclusion Criteria

+ Women 15-44 years old
+ Delivery of an infant >=20 weeks between 2007-2020
+ Continuous enrollment in TennCare from 90 pre-delivery through 41 days post partum
+ No sickle cell disease or active malignancy
+ With and without MOUD receipt initially, but only included those with MOUD for main analysis
+ Up to 5-days gap of continuous enrollment

<br>

+ Exposure: Days of MOUD supply between -90 days prior to delivery and +41 days post-delivery (maximum number of days = 132)
+ Covariates also collected during exposure period 
+ Outcomes: first event of either overdose or death from 42 days post-delivery

<br>

Statistical analysis

Descriptive statistics were computed using frequencies (proportions) and medians (IQRs) as appropriate for the overall sample and also stratified by timing of MOUD receipt. Timing of MOUD receipt variable was categorized relative to timing of delivery and has two levels: receipt during one period only either pre- or post-delivery or receipt during both pre- and post-delivery periods. Due to the low number of outcomes, several variables were collapsed into one. These variables are listed below. 

<br>

+ Route of delivery: Cesarean or Vaginal (Vaginal + Forcep + Vacuum)
+ Pain conditions: Back pain + dental pain + fibromyalgia + indicator for NSAIDs 
+ Alcohol use disorder + other substance use
+ Benzodiazepine + Anticonvulsant use 
+ Anxiety + Depression

<br>

After calculating the number of days of MOUD receipt, various histograms were plotted to examine differences by method. Then, multivariable Cox Proportional hazards model were used to determine the associations between total days of MOUD receipt as the main exposure and death or overdose as the outcome. For individuals with multiple outcomes, the earliest outcome was used. Outcomes that occurred during the follow-up period were counted if they were captured in the TN all payer database if an individual lost eligibility for TennCare. Average dose per day of MOUD receipt (qty*dose)/(days supply) was also calculated and described in the table. Average days supply and dose from the pre-delivery period that exceeded 90 days was carried over and added to the post-delivery period. 

<br>

The Cox model was first modeled using linear continuous terms, due to the limited degrees of freedom. However, year of delivery was determined to be non-linear and was modeled using a restricted cubic spline with 3 knots. These models included the following covariates: year of delivery, total day MOUD receipt, age at delivery, gestation weeks, route of delivery, combined pain conditions, alcohol/other substance use, benzodiazepine/anticonvulsant use, anxiety/depression, maternal morbidity (CDC defined). Additional covariates were included initially but were removed due to the small number of outcomes. These include race/ethnicity combined, number of previous pregnancies, schizophrenia/personality disorder, CDC social vulnerability index, and timing of MOUD receipt (one or two periods relative to delivery). Robust standard errors clustered at the individual level were computed. 


Additonal notes: few outcomes (<5) prior to 2015


<br>

# Descriptive Statistics (90 days pre-delivery & up to 5 days loss enroll)
## Demographics/Clinical Characteristics
### Column percents

<br>

<!-- * There are 13941 pregnancies in 11635 unique individuals.  -->

* There are 15103 pregnancies in 12532 unique individuals

```{r}
label(basic$total.rx.days.exp) <- "Total days supplied MOUD"
label(basic$total.rx.days.total_3bup) <- "Total days supplied buprenorphine"
moud <- basic[basic$moud.yes==1, ]
# save moud data
# write.csv(moud, file="moud_cohort_202601.csv", row.names = FALSE)
# save(moud, file="moud_cohorot_202601.Rdata")


# add later
# death.overdose.new.f +  death.from.od.new.f
# add final.outcome.f later
tangram(paste('moud.yes.f~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
          final.outcome.f + final.outcome2.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
          total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
          total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] + avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] +
          total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] + rxtype +
          pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
              noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
        data=basic,  pformat=4, digits=2, style="nejm", test=TRUE)



# tangram(paste('moud.yes.f~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
#           final.outcome.f + final.outcome2.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
#           total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
#           pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
#               noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
#         data=basic,  pformat=4, digits=2, style="nejm", test=TRUE)


# tangram(moud.yes.f~ delivery_year, data =basic)

# tangram(moud.yes.f~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
# final.outcome.f + final.outcome2.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
# total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
# total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] + avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] +
# total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] +
# pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f, 
# data=basic,  pformat=4, digits=2, style="nejm", test=TRUE)

```

<br>

* Outlier with average dose of 16,000 due to quantity of (56,000 * 8mg)/28 days 

<!-- ### MOUD information -->

<!-- ```{r} -->
<!-- tangram(1~ moud.pre + moud.pre2 + moud.pre3 + moud.post + moud.post2 + moud.post3 + -->
<!--           moud.both + moud.both2 + moud.both3, data= basic, pformat=4, digits=2, style= "nejm", collapse_single=FALSE) -->
<!-- ``` -->
<!-- <br> -->


### Characteristics based on MOUD timing (method 3)
```{r}
tangram(paste('moud.both3div~ delivery_year[0] + delivery.year.cat + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
          final.outcome.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
          total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
          total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] +  avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] +
          total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] + rxtype+
          pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
              noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
        data=basic, pformat=4, digits=2, style="nejm")
```

<br>
  
## Histogram of MOUD days supplied
### Before delivery up to 90 days (Method 1)
```{r, warning=FALSE, message=FALSE, comment=FALSE}
ggplot(basic, aes(x=total.rx.days.b90)) + 
  geom_histogram(color="black", fill="white", binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 90, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

### Before delivery up to 90 days (Method 2)
```{r}
ggplot(basic, aes(x=total.rx.days.b90_2)) + 
  geom_histogram(color="black", fill="white",  binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 90, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>
  
### Before delivery up to 90 days (Method 3)
```{r}
ggplot(basic, aes(x=total.rx.days.b90_3)) + 
  geom_histogram(color="black", fill="white",  binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 90, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```


<br>
  
### After delivery up to 41 days (Method 1)
```{r}
ggplot(moud, aes(x=total.rx.days.p41_2)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 30, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

### After delivery up to 41 days (Method 2)
```{r}
ggplot(moud, aes(x=total.rx.days.p41_2)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 30, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>
  
### After delivery up to 41 days (Method 3)
```{r}
ggplot(basic, aes(x=total.rx.days.p41_3)) + 
  geom_histogram(color="black", fill="white", binwidth = 3) +
  scale_x_continuous(breaks = seq(0, 40, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```


<br>
  
### Entire exposure (Method 1)
```{r}
ggplot(basic, aes(x=total.rx.days.exp)) + 
  geom_histogram(color="black", fill="white", binwidth=10) +
  scale_x_continuous(breaks = seq(0, 130, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>
  
  
### Entire exposure (Method 2)
```{r}
ggplot(basic, aes(x=total.rx.days.exp2)) + 
  geom_histogram(color="black", fill="white", binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 130, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>
  
### Entire exposure (Method 3)
```{r}
ggplot(basic, aes(x=total.rx.days.exp3)) + 
  geom_histogram(color="black", fill="white", binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 130, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```
<br>

### Entire exposure buprenorphine only (Method 3)
```{r}
ggplot(basic, aes(x=total.rx.days.total_3bup)) + 
  geom_histogram(color="black", fill="white", binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 130, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```


<br>
  
  
# Survival Analysis
## Kaplan Meier Plot by MOUD
```{r}
  kmout <- survfit(Surv(days.to.event, outcome.bin) ~moud.yes, data=basic)
  # test with survminer
  # survminer::ggsurvplot(kmout, data=basic, fun="event")

  #KM survival curve for only 1 year
  plot(kmout, conf.int=FALSE, col=c("red", "blue"), xlab="Days Since 41-days Post-Partum", ylab="Probability of Death or Overdose", main= "KM curves by MOUD status", xlim=c(0, 323), fun="event")
  legend('bottomright', c("No MOUD", "MOUD"), col=c("red", "blue"), lty=1)
  
  # 
  # 
  #  kmout2 <- survfit(Surv(days.to.event2, outcome.bin) ~moud.yes, data=basic)
  # # test with survminer
  # # survminer::ggsurvplot(kmout, data=basic, fun="event")
  # 
  # #KM survival curve for only 1 year
  # plot(kmout2, conf.int=FALSE, col=c("red", "blue"), xlab="Days Since 41-days Post-Partum", ylab="Probability of Death or Overdose", main= "KM curves by MOUD status", xlim=c(0, 323), fun="event")
  # legend('bottomright', c("No MOUD", "MOUD"), col=c("red", "blue"), lty=1)
```

  <br>

## Log-Rank Test

```{r}
  surv_diff <- survdiff(Surv(days.to.event, outcome.bin) ~moud.yes, data=basic)
  print(surv_diff)

  # 
  # surv_diff2 <- survdiff(Surv(days.to.event2, outcome.bin) ~moud.yes, data=basic)
  # print(surv_diff2)

```
  * There is a statistically significant difference in death or overdose between the two groups

  <br>

<!-- ## Kaplan Meier Plot one group -->
<!-- ```{r} -->
<!--   kmout1 <- survfit(Surv( death.or.od) ~1, data=basic) -->

<!--   #KM survival curve for only 1 year -->
<!--   plot(kmout1, conf.int=TRUE, xlab="Days from Delivery", ylab="Probability of Death or Overdose", main= "KM curve overall",  xlim=c(0, 365), fun="event") -->
<!-- ``` -->
  
## Multivariable Cox PH model
### Linear continuous terms
```{r, warning=FALSE, message=NA, include=FALSE}

dd <- datadist(basic)
options(datadist= 'dd')
dd$limits$pain.combine.f[2] <- "No"
dd$limits$alcuse.othersub.f[2] <- "No"
dd$limits$delivery_year[2] <- 2014
dd$limits$rpl_themes[2] <- 0.25
# 
# #follow up time should be log with MI interacted with outcome for MI
# set.seed(90)
# 
# basic$follow.up.days.log10 <- log10(basic$follow.up.days + 0.01)
# basic$follow.up.days.minus42 <- basic$follow.up.days - 42
# basic$follow.up.days.minus42.log10 <- log10(basic$follow.up.days.minus42 + 0.01)
# 
# # interact followup days with outcome # standard procedure in biostats
# 
# #Covariates are age, race/ethnicity combined, SVI, route of delivery, maternal morbidity, number of prior pregnancies, gestational weeks, number of MOUD days, pain combined, personality disorders combined, anxiety, depression, substance use combined, medications combined
# 
# 
# cph_multi_i <- aregImpute(~ (follow.up.days.log10*death.or.od) + momage + momraceeth.new + rpl_themes + routedeliv.new + previouspregnanciestotaln + gweeks + total.rx.days.exp +
#                             pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f , data=basic, n.impute = 20)
# 
# cph_multi <- fit.mult.impute(Surv(follow.up.days.log10, death.or.od) ~ moud.yes.f + total.rx.days.exp + delivery_year+  momage + momraceeth_comb.new + rpl_themes + routedeliv.new + previouspregnanciestotaln + gweeks +
#                             pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f,
#                              data=basic, fitter = cph, xtrans= cph_multi_i, n.impute=20, y=TRUE, x=TRUE, surv=TRUE)

cph_multi <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.exp3 +  momage + momraceeth_comb.new + routedeliv.new + gweeks +  
                   pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f + rpl_themes +
                   previouspregnanciestotaln + strat(moud.yes) , data=basic, x=TRUE, y=TRUE) 

# cph_multi <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.exp3 +  momage + routedeliv.new + gweeks +  
#                    pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f + moud.yes.f, data=basic, x=TRUE, y=TRUE) 
# cox.zph(cph_multi) # PH assumption met
```

```{r}
print(summary(robcov(cph_multi, cluster = basic$unique_id.x), momage=c(27, 31), total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), rpl_themes=c(0.25, 0.75),  vnames = 'labels'), digits=2)
print(anova(robcov(cph_multi), vnames="labels"), digits=3)

```

```{r, fig.height=6, fig.width=8}
plot(summary(cph_multi, momage=c(27, 31),  total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39, 37), vnames = 'labels'))
```

```{r, fig.height=12, fig.width=12}
ggplot(Predict(cph_multi, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical")
```  

<br>

### With splines
```{r, include=FALSE, warning=FALSE, message=FALSE}

cph_multi_sp <- cph(Surv(days.to.event, outcome.bin) ~ rcs(delivery_year,3) + rcs(total.rx.days.exp3,3) +  momage + momraceeth_comb.new + routedeliv.new + gweeks +  
                   pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f + rpl_themes +
                   previouspregnanciestotaln + strat(moud.yes) , data=basic, x=TRUE, y=TRUE) 

```

```{r}
# print(summary(robcov(cph_multi_sp, cluster = basic$unique_id.x), delivery_year=c(2016, 2020),  momage=c(25, 30), rpl_themes=c(0.5, 0.75), total.rx.days.exp=c(0,1), vnames = 'labels'), digits=3)
# print(anova(robcov(cph_multi_sp), vnames="labels"), digits=3)
# x <- Predict(cph_multi_sp, delivery_year=c(2007:2020))
# x$yhat_exp <- exp(x$yhat)
# x$lower_exp <- exp(x$lower)
# x$upper_exp <- exp(x$upper)

print(summary(robcov(cph_multi_sp, cluster = basic$unique_id.x), momage=c(27, 31), total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), rpl_themes=c(0.25, 0.75), gweeks=c(39, 37),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2007,2015),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2010,2014),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2015, 2017),  vnames = 'labels'))
# 
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2015, 2020),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2017, 2020),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2017, 2018),  vnames = 'labels'))

print(anova(robcov(cph_multi_sp), vnames="labels"), digits=3)

```

```{r,fig.height=6, fig.width=8}
plot(summary(cph_multi_sp, momage=c(27, 31),  delivery_year=c(2016, 2018), gweeks=c(39, 37), total.rx.days.exp3=c(1, 31),rpl_themes=c(0.25, 0.75),  vnames = 'labels'))
```

```{r, fig.height=12, fig.width=12}
ggplot(Predict(cph_multi_sp, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical")

```

<br>

# Propensity Score (Matching Weights) by MOUD receipt

```{r}
# variables need to be numeric for PSW later
basic$momraceeth_comb.new.num <- as.numeric(basic$momraceeth_comb.new)
basic$routedeliv.new.num <- as.numeric(basic$routedeliv.new)

# number of previous pregnancies and SVI removed due to missing values

form.ps <- "moud.yes ~ delivery_year +  momage + momraceeth_comb.new.num + routedeliv.new.num + gweeks + pain.combine + personality.combine + alcuse.othersub + benzo.anticonv + anxiety + depression + mat_morbidity"

V.name <- c("delivery_year", "momage" , "momraceeth_comb.new.num" , "routedeliv.new.num" , "gweeks", "pain.combine", "personality.combine", "alcuse.othersub", "benzo.anticonv", "anxiety", "depression", "mat_morbidity")


mw <- psw(data=basic, form.ps = form.ps, weight="MW", mirror.hist = TRUE, V.name = V.name, add.weight = TRUE)  

ps1balance <- psw.balance(data= basic, weight="MW", form=form.ps, V.name = V.name)


ps1balancebefore <- as.data.frame(ps1balance$std.diff.before)
ps1balancebefore$std.diff.pct.prop <- round(abs(ps1balancebefore$std.diff.pct)/100, digits=2)
ps1balanceafter <- as.data.frame(ps1balance$std.diff.after)
ps1balanceafter$std.diff.pct.prop <- round(abs(ps1balanceafter$std.diff.pct)/100, digits=2)
ps1balance.to.plot <- data.frame(var= rownames(ps1balancebefore), Before= ps1balancebefore$std.diff.pct.prop, After= ps1balanceafter$std.diff.pct.prop)
ps1balance.to.plot$Variable <- c("Delivery year", "Age", "Race/ethnicity", "Route of delivery", "Gestational age", "Pain condition", "Personality disorder", "Alc/substance use", "Benzo/anticonvulsants", "Anxiety", "Depression", "Maternal morbidity")

ps1balance.to.plot <- melt(ps1balance.to.plot, id="Variable")
ps1balance.to.plot <- subset(ps1balance.to.plot, variable!= "var")
ps1balance.to.plot$value <- as.numeric(ps1balance.to.plot$value)
ps1balance.to.plot$Variable <- with(ps1balance.to.plot, reorder(x= Variable, X= value, FUN= sum))
```
<br>


* Mirror histograms show propensity scores of the control group above the horizontal line in white, while the treatment group is below the horizontal line in dark green. Light green bars display the summation of weights of subjects within each propensity score stratum. This plot shows good overlap between the control and treatment groups, and the close similarity between them after applying matching weights. 



## Absolute standardized mean difference plot
```{r}
ggplot(ps1balance.to.plot, aes(x=value, y=Variable), fill=variable) +
  geom_point(aes(color= variable), size=3) + labs(x= "SMD", y="") + theme_minimal() + theme(legend.title = element_blank())

#grayscale figure
# ggplot(smd1.to.plot, aes(x=value, y=Variable), fill=variable) +
#   geom_point(aes(color= variable), size=3) +
#   scale_colour_grey()+
#   labs(x= "SMD", y="") +
#   theme_minimal() + theme(legend.title = element_blank(), text = element_text(size=14))

# graph2ppt(file="smd_plot.pptx")
```
<br>

## Logistic regression model
```{r, warning=FALSE}
pslog1 <- glm(outcome.bin ~ moud.yes, family=binomial(link="logit"), weights = mw$W, data=basic)

summary(pslog1)
```

<br>

* exponentiated Odds Ratio
```{r, warning=FALSE}
exp(cbind(OR = coef(pslog1), confint(pslog1)))
```

<br>

<!-- # Propensity score excluding naltrexone/methadone by MOUD receipt -->


<!-- ```{r} -->
<!-- # variables need to be numeric for PSW later -->
<!-- basic2 <- basic[!(basic$last_rx_type %in% c("Methadone", "Naltrexone", "Naltrexone Hydrochloride", "Naltrexone, Naltrexone Hydrochloride")), ] -->
<!-- basic2$last_rx_type <- droplevels(basic2$last_rx_type) -->

<!-- # number of previous pregnancies and SVI removed due to missing values -->

<!-- form.ps <- "moud.yes ~ delivery_year +  momage + momraceeth_comb.new.num + routedeliv.new.num + gweeks + pain.combine + personality.combine + alcuse.othersub + benzo.anticonv + anxiety + depression + mat_morbidity" -->

<!-- V.name <- c("delivery_year", "momage" , "momraceeth_comb.new.num" , "routedeliv.new.num" , "gweeks", "pain.combine", "personality.combine", "alcuse.othersub", "benzo.anticonv", "anxiety", "depression", "mat_morbidity") -->


<!-- mw <- psw(data=basic2, form.ps = form.ps, weight="MW", mirror.hist = TRUE, V.name = V.name, add.weight = TRUE)   -->

<!-- ps1balance <- psw.balance(data= basic2, weight="MW", form=form.ps, V.name = V.name) -->


<!-- ps1balancebefore <- as.data.frame(ps1balance$std.diff.before) -->
<!-- ps1balancebefore$std.diff.pct.prop <- round(abs(ps1balancebefore$std.diff.pct)/100, digits=2) -->
<!-- ps1balanceafter <- as.data.frame(ps1balance$std.diff.after) -->
<!-- ps1balanceafter$std.diff.pct.prop <- round(abs(ps1balanceafter$std.diff.pct)/100, digits=2) -->
<!-- ps1balance.to.plot <- data.frame(var= rownames(ps1balancebefore), Before= ps1balancebefore$std.diff.pct.prop, After= ps1balanceafter$std.diff.pct.prop) -->
<!-- ps1balance.to.plot$Variable <- c("Delivery year", "Age", "Race/ethnicity", "Route of delivery", "Gestational age", "Pain condition", "Personality disorder", "Alc/substance use", "Benzo/anticonvulsants", "Anxiety", "Depression", "Maternal morbidity") -->

<!-- ps1balance.to.plot <- melt(ps1balance.to.plot, id="Variable") -->
<!-- ps1balance.to.plot <- subset(ps1balance.to.plot, variable!= "var") -->
<!-- ps1balance.to.plot$value <- as.numeric(ps1balance.to.plot$value) -->
<!-- ps1balance.to.plot$Variable <- with(ps1balance.to.plot, reorder(x= Variable, X= value, FUN= sum)) -->
<!-- ``` -->



<!-- * Mirror histograms show propensity scores of the control group above the horizontal line in white, while the treatment group is below the horizontal line in dark green. Light green bars display the summation of weights of subjects within each propensity score stratum. This plot shows good overlap between the control and treatment groups, and the close similarity between them after applying matching weights.  -->



<!-- ## Absolute standardized mean difference plot -->
<!-- ```{r} -->
<!-- ggplot(ps1balance.to.plot, aes(x=value, y=Variable), fill=variable) + -->
<!--   geom_point(aes(color= variable), size=3) + labs(x= "SMD", y="") + theme_minimal() + theme(legend.title = element_blank()) -->

<!-- #grayscale figure -->
<!-- # ggplot(smd1.to.plot, aes(x=value, y=Variable), fill=variable) + -->
<!-- #   geom_point(aes(color= variable), size=3) + -->
<!-- #   scale_colour_grey()+ -->
<!-- #   labs(x= "SMD", y="") + -->
<!-- #   theme_minimal() + theme(legend.title = element_blank(), text = element_text(size=14)) -->

<!-- # graph2ppt(file="smd_plot.pptx") -->
<!-- ``` -->


<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog1 <- glm(outcome.bin ~ moud.yes, family=binomial(link="logit"), weights = mw$W, data=basic2) -->

<!-- summary(pslog1) -->
<!-- ``` -->

<!-- <br> -->


<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog1), confint(pslog1))) -->
<!-- ``` -->
<!-- <br> -->


<!-- # Propensity score excluding naltrexone/methadone by MOUD receipt (2015+) -->

<!-- ```{r} -->
<!-- # variables need to be numeric for PSW later -->
<!-- basic2015 <-basic2[basic2$delivery_year>=2015, ] -->

<!-- # number of previous pregnancies and SVI removed due to missing values -->

<!-- form.ps2015 <- "moud.yes ~ delivery_year +  momage + momraceeth_comb.new.num + routedeliv.new.num + gweeks + pain.combine + personality.combine + alcuse.othersub + benzo.anticonv + anxiety + depression + mat_morbidity" -->

<!-- V.name2015 <- c("delivery_year", "momage" , "momraceeth_comb.new.num" , "routedeliv.new.num" , "gweeks", "pain.combine", "personality.combine", "alcuse.othersub", "benzo.anticonv", "anxiety", "depression", "mat_morbidity") -->


<!-- mw <- psw(data=basic2015, form.ps = form.ps2015, weight="MW", mirror.hist = TRUE, V.name = V.name2015, add.weight = TRUE)   -->

<!-- ps1balance2015 <- psw.balance(data= basic2015, weight="MW", form=form.ps, V.name = V.name) -->


<!-- ps1balancebefore2015 <- as.data.frame(ps1balance2015$std.diff.before) -->
<!-- ps1balancebefore2015$std.diff.pct.prop <- round(abs(ps1balancebefore2015$std.diff.pct)/100, digits=2) -->
<!-- ps1balanceafter <- as.data.frame(ps1balance$std.diff.after) -->
<!-- ps1balanceafter$std.diff.pct.prop <- round(abs(ps1balanceafter$std.diff.pct)/100, digits=2) -->
<!-- ps1balance.to.plot2015 <- data.frame(var= rownames(ps1balancebefore2015), Before= ps1balancebefore2015$std.diff.pct.prop, After= ps1balanceafter$std.diff.pct.prop) -->
<!-- ps1balance.to.plot2015$Variable <- c("Delivery year", "Age", "Race/ethnicity", "Route of delivery", "Gestational age", "Pain condition", "Personality disorder", "Alc/substance use", "Benzo/anticonvulsants", "Anxiety", "Depression", "Maternal morbidity") -->

<!-- ps1balance.to.plot2015 <- melt(ps1balance.to.plot2015, id="Variable") -->
<!-- ps1balance.to.plot2015 <- subset(ps1balance.to.plot2015, variable!= "var") -->
<!-- ps1balance.to.plot2015$value <- as.numeric(ps1balance.to.plot2015$value) -->
<!-- ps1balance.to.plot2015$Variable <- with(ps1balance.to.plot2015, reorder(x= Variable, X= value, FUN= sum)) -->
<!-- ``` -->



<!-- * Mirror histograms show propensity scores of the control group above the horizontal line in white, while the treatment group is below the horizontal line in dark green. Light green bars display the summation of weights of subjects within each propensity score stratum. This plot shows good overlap between the control and treatment groups, and the close similarity between them after applying matching weights.  -->



<!-- ## Absolute standardized mean difference plot -->
<!-- ```{r} -->
<!-- ggplot(ps1balance.to.plot2015, aes(x=value, y=Variable), fill=variable) + -->
<!--   geom_point(aes(color= variable), size=3) + labs(x= "SMD", y="") + theme_minimal() + theme(legend.title = element_blank()) -->

<!-- ``` -->


<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog1.2015 <- glm(outcome.bin ~ moud.yes, family=binomial(link="logit"), weights = mw$W, data=basic2015) -->

<!-- summary(pslog1.2015) -->
<!-- ``` -->

<!-- <br> -->


<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog1.2015), confint(pslog1.2015))) -->
<!-- ``` -->


# Only including data of women with MOUD receipt 
## Descriptive Statistics

<br>

<!-- * There are 7055 pregnancies in 6049 unique individuals.  -->

* There are 7472 pregnancies in 6387 unique individuals
```{r}

# add later
# death.overdose.new.f +  death.from.od.new.f
# add final.outcome.f later
tangram(paste('1~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
          final.outcome.f + final.outcome2.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
          total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
          total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] + avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] +
          total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] + rxtype+
          pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
            noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
        data=moud,  pformat=4, digits=2, style="nejm")
```
<br>

## By cases
```{r}
moud$outcome.bin.f <- factor(moud$outcome.bin, levels= 0:1, labels=c("No outcome", "Had outcome"))
label(moud$outcome.bin.f) <- "Outcome"  
  
tangram(paste('outcome.bin.f~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
          final.outcome.f + final.outcome2.f + 
          total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] + avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] +
          total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] +
          pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
            noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
        data=moud,  pformat=4, digits=2, style="nejm", test=TRUE)
```


<br>

## MOUD information

```{r}
tangram(1~ moud.pre + moud.pre2 + moud.pre3 + moud.post + moud.post2 + moud.post3 +
          moud.both + moud.both2 + moud.both3, data= moud, pformat=4, digits=2, style= "nejm", collapse_single=FALSE)
```
<br>



## Characteristics based on MOUD timing (method 3)
### Two categories
```{r}
moud$moud.both3 <- droplevels(moud$moud.both3)
label(moud$moud.both3) <- "MOUD receipt timing"

moud$final.outcome.3lev <- ifelse(moud$final.outcome==1, 1, 
                                  ifelse(moud$final.outcome==2, 2, NA))
moud$final.outcome.3lev <- ifelse(is.na(moud$final.outcome.3lev), 3, moud$final.outcome.3lev)
moud$final.outcome.3lev.f <- factor(moud$final.outcome.3lev, levels=1:3, labels=c("Death", "Overdose", "Censored"))
label(moud$final.outcome.3lev.f) <- "Final outcome 3 levels"

tangram(paste('moud.both3~ delivery_year[0] + delivery.year.cat + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
          final.outcome.f + final.outcome.3lev.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
          total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
          total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] +  avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] + 
          total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] +
          pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
            noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
        data=moud, test=TRUE, pformat=4, digits=2, style="nejm")
```

<br>

### Three categories
```{r}
moud$moud.both3div <- droplevels(moud$moud.both3div)
label(moud$moud.both3div) <- "MOUD receipt timing"
tangram(paste('moud.both3div~ delivery_year[0] + delivery.year.cat + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
          final.outcome.f + final.outcome.3lev.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
          total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
          total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] +  avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] + 
          total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] +
          pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
            noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
        data=moud, test=TRUE, pformat=4, digits=2, style="nejm")
```



<!-- ### Row Percents -->
<!-- ```{r} -->
<!-- tangram(paste('moud.yes.f~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] + -->
<!--           death.new.f + death.or.od.f + death.or.op.od + final.outcome.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] +  -->
<!--           pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f +', -->
<!--             noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'),  -->
<!--         data=basic,  overall=TRUE, test=TRUE, pformat=4, digits=2, style="nejm", row_percents=TRUE) -->
<!-- ``` -->

<br>

## Histogram of MOUD days supplied
### Before delivery up to 90 days (Method 1)
```{r}
ggplot(moud, aes(x=total.rx.days.b90)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 90, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

### Before delivery up to 90 days (Method 2)
```{r}
ggplot(moud, aes(x=total.rx.days.b90_2)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 90, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>

### Before delivery up to 90 days (Method 3)
```{r}
ggplot(moud, aes(x=total.rx.days.b90_3)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 90, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>

### After delivery up to 41 days (Method 1)
```{r}
ggplot(moud, aes(x=total.rx.days.p41_2)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 30, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

### After delivery up to 41 days (Method 3)
```{r}
ggplot(moud, aes(x=total.rx.days.p41_2)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 30, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>

### After delivery up to 41 days (Method 3)
```{r}
ggplot(moud, aes(x=total.rx.days.p41_3)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 30, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```


<br>

### Entire exposure (Method 1)
```{r}
ggplot(moud, aes(x=total.rx.days.exp)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 130, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>


### Entire exposure (Method 2)
```{r}
ggplot(moud, aes(x=total.rx.days.exp2)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 130, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>

### Entire exposure (Method 3)
```{r}
ggplot(moud, aes(x=total.rx.days.exp3)) + 
  geom_histogram(color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 130, 10)) + 
  labs(x= "Days", y="Count") +
  theme_classic()
```

<br>

<!-- ### Histogram by MOUD post-delivery group -->
<!-- ```{r} -->
<!-- moud %>% -->
<!--   ggplot(aes(x=total.rx.days.exp3, fill=moud.both3)) +  -->
<!--   geom_histogram(alpha=0.75, color=1) + -->
<!--   labs(x= "Days of MOUD supply", y="Count of pregnancies") + -->
<!--   scale_color_manual(values=c("black", "tan3")) + -->
<!--   scale_fill_manual(values=c("black", "tan3")) + -->
<!--   guides(fill= guide_legend(title= "Receipt of MOUD relative to delivery")) + -->
<!--   theme_classic() + -->
<!--   theme(legend.box = "vertical", legend.position = "top")  -->


<!-- #  -->
<!-- # moud %>% -->
<!-- #   ggplot(aes(x=total.rx.days.exp3, color=moud.both3)) +  -->
<!-- #   geom_histogram() + -->
<!-- #   labs(x= "Days of MOUD supply", y="Count of individuals", fill="Receipt of MOUD relative to delivery") + -->
<!-- #   scale_color_manual(values=c("lightgray", "darkgray")) + -->
<!-- #   theme_classic() + -->
<!-- #   theme(legend.box = "vertical", legend.position = "top")  -->
<!-- #  -->
<!-- # ggplot(moud,aes(x=total.rx.days.exp3))+geom_histogram()+facet_grid(~moud.both3)+theme_bw() -->
<!-- ``` -->
<!-- <br> -->


<!-- ### Histogram by MOUD post-delivery group -->
<!-- ```{r} -->
<!-- moud %>% -->
<!--   ggplot(aes(x=total.rx.days.exp3, fill=moud.both3div)) +  -->
<!--   geom_histogram(alpha=0.75, color=1) + -->
<!--   labs(x= "Days of MOUD supply", y="Count of pregnancies") + -->
<!--   scale_color_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   scale_fill_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   guides(fill= guide_legend(title= "Receipt of MOUD relative to delivery")) + -->
<!--   theme_classic() + -->
<!--   theme(legend.box = "vertical", legend.position = "top")  -->

<!-- ``` -->

<!-- <br> -->

<!-- ```{r, fig.width=10, fig.height= 5} -->
<!-- custom_x <- list( -->
<!--   scale_x_continuous(limits=c(0, 90)), -->
<!--   scale_x_continuous(limits=c(0, 42)), -->
<!--   scale_x_continuous(limits=c(0, 132)) -->
<!-- ) -->

<!-- sample_size_labs <- c( -->
<!--   `Only pre` = "Only pre (n=705)", -->
<!--   `Only post` = "Only post (n=1158)", -->
<!--   `Both pre and post` = "Both pre and post (n=1940)" -->
<!-- ) -->

<!-- moud %>% -->
<!--   ggplot(aes(x=total.rx.days.exp3, fill=moud.both3div)) +  -->
<!--   geom_histogram(alpha=0.75, color=1) +  -->
<!--   facet_wrap(~moud.both3div, scales = "free_x") + -->
<!--   facetted_pos_scales(x=custom_x) + -->
<!--   labs(x= "Days of MOUD supply", y="Count of pregnancies") + -->
<!--   scale_color_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   scale_fill_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   # guides(fill= guide_legend(title= "Receipt of MOUD relative to delivery")) + -->
<!--   theme_classic() + -->
<!--   theme(legend.box = "vertical", legend.position = "top", text = element_text(size=18))  -->


<!-- moud %>% -->
<!--   ggplot(aes(x=total.rx.days.exp3, fill=moud.both3div)) +  -->
<!--   geom_histogram(alpha=0.75, color = 1) +  -->
<!--   facet_wrap(~moud.both3div, scales = "free_x", labeller = as_labeller(sample_size_labs)) + -->
<!--   facetted_pos_scales(x=custom_x) + -->
<!--   labs(x= "Days of MOUD supply", y="Count of pregnancies", title = "Receipt of MOUD relative to delivery") + -->
<!--   scale_color_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   scale_fill_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   guides(fill= guide_legend(title= "Receipt of MOUD relative to delivery")) + -->
<!--   theme_classic() + -->
<!--   theme(legend.box = "vertical", legend.position = "none", text = element_text(size=18), plot.title = element_text(hjust = 0.5))  -->


<!-- ``` -->



<br>

# Survival Analysis {-}

## Kaplan Meier Plot 
```{r}
kmoutm <- survfit(Surv(days.to.event, outcome.bin) ~1, data=moud)
print(kmoutm)
# test with survminer
# survminer::ggsurvplot(kmout, data=basic, fun="event")

#KM survival curve for only 1 year
# plot(kmoutm, conf.int=FALSE, col=c("red", "blue"), xlab="Days since 41-days post-partum", ylab="Probability of Death or Overdose", main= "KM curves by MOUD timing", xlim=c(0, 365), fun="event")
# legend('bottomright', c("Either pre or post", "Both pre and post"), col=c("red", "blue"), lty=1)

plot(kmoutm, conf.int=FALSE, col=c("red", "blue"), xlab="Days since 41-days post-partum", ylab="Probability of Death or Overdose", xlim=c(0, 323), fun="event")

summary(kmoutm$time)

kable(data.frame(kmoutm$time, kmoutm$surv), col.names = c("time", "probability"))
```

<br>

## Time to event for overdose and death
```{r}
deathod <- moud[which(moud$final.outcome.f %in% c("Death", "Overdose")), ]
describe(deathod$days.to.event)

death <- deathod[which(deathod$final.outcome.f=="Death"), ]
describe(death$days.to.event)

od <- deathod[which(deathod$final.outcome.f=="Overdose"), ]
describe(od$days.to.event)
```


<!-- ## Log-Rank Test -->

<!-- ```{r} -->
<!-- surv_diffm <- survdiff(Surv(days.to.event, outcome.bin) ~moud.both3, data=moud) -->
<!-- print(surv_diffm) -->

<!-- ``` -->
<!-- * There is no statistically significant difference in death or overdose between the two groups -->

<!-- <br> -->

<!-- ## Kaplan Meier Plot one group -->
<!-- ```{r} -->
<!-- kmout1 <- survfit(Surv(follow.up.days, death.or.od) ~1, data=basic) -->

<!-- #KM survival curve for only 1 year -->
<!-- plot(kmout1, conf.int=TRUE, xlab="Days from Delivery", ylab="Probability of Death or Overdose", main= "KM curve overall",  xlim=c(0, 365), fun="event") -->
<!-- ``` -->

# Multivariable Cox PH model
## Linear continuous terms
```{r, warning=FALSE, message=NA, include=FALSE}

dd <- datadist(moud)
options(datadist= 'dd')
dd$limits$pain.combine.f[2] <- "No"
dd$limits$alcuse.othersub.f[2] <- "No"
dd$limits$moud.both3[2] <- "Either pre or post"
# 
# #follow up time should be log with MI interacted with outcome for MI
# set.seed(90)
# 
# basic$follow.up.days.log10 <- log10(basic$follow.up.days + 0.01)
# basic$follow.up.days.minus42 <- basic$follow.up.days - 42
# basic$follow.up.days.minus42.log10 <- log10(basic$follow.up.days.minus42 + 0.01)
# 
# # interact followup days with outcome # standard procedure in biostats
# 
# #Covariates are age, race/ethnicity combined, SVI, route of delivery, maternal morbidity, number of prior pregnancies, gestational weeks, number of MOUD days, pain combined, personality disorders combined, anxiety, depression, substance use combined, medications combined
# 
# 
# cph_multi_i <- aregImpute(~ (follow.up.days.log10*death.or.od) + momage + momraceeth.new + rpl_themes + routedeliv.new + previouspregnanciestotaln + gweeks + total.rx.days.exp +
#                             pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f , data=basic, n.impute = 20)
# 
# cph_multi <- fit.mult.impute(Surv(follow.up.days.log10, death.or.od) ~ moud.yes.f + total.rx.days.exp + delivery_year+  momage + momraceeth_comb.new + rpl_themes + routedeliv.new + previouspregnanciestotaln + gweeks +
#                             pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f,
#                              data=basic, fitter = cph, xtrans= cph_multi_i, n.impute=20, y=TRUE, x=TRUE, surv=TRUE)

cph_multim <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.exp3 +  momage +  routedeliv.new + gweeks +  
                   pain.combine.f +  alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE) 
# cox.zph(cph_multim) # PH assumption met
```

```{r}

print(summary(robcov(cph_multim, cluster = moud$unique_id.x), momage=c(27,31), total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39, 37),
              vnames = 'labels'))
print(anova(robcov(cph_multim), vnames="labels"), digits=2)

```

```{r, fig.height=6, fig.width=8}
plot(summary(cph_multim, momage=c(27, 31), total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39, 37),  vnames = 'labels'))
```

```{r, fig.height=12, fig.width=12}
ggplot(Predict(cph_multim, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical")
# ggplot(Predict(cph_multim, delivery_year=c(2007:2020), fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical")
probdf <- Predict(cph_multim, total.rx.days.exp3=c(1, 31, 61, 89, 119, 132), delivery_year=c(2007:2020), momage=27, gweeks=39, fun=plogis)

ggplot(Predict(cph_multim, delivery_year=c(2007:2020), total.rx.days.exp3=89, momage=27, gweeks=39, fun=plogis), ylab="Probability of Outcome") + scale_x_continuous(breaks = c(2008, 2010, 2012, 2014, 2016, 2018, 2020))
ggplot(Predict(cph_multim, total.rx.days.exp3=c(1:132), delivery_year=2016, momage=27, gweeks=39, fun=plogis), ylab="Probability of Outcome", xlab = "Total Days") + scale_x_continuous(breaks = c(0, 30, 60, 90, 120))
```

<br>

### Univariable HR
```{r}
cph_uni <- cph(Surv(days.to.event, outcome.bin) ~ total.rx.days.exp3 , data=moud, x=TRUE, y=TRUE) 

print(summary(robcov(cph_uni, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), vnames = 'labels'))
print(anova(robcov(cph_uni), vnames="labels"), digits=2)
```



## With splines
```{r, include=FALSE, warning=FALSE, message=FALSE}
# cph_multi_sp <- fit.mult.impute(Surv(follow.up.days.log10, death.or.od) ~  rcs(delivery_year, 3)  + total.rx.days.exp  + momage + momraceeth_comb.new + rpl_themes + routedeliv.new + previouspregnanciestotaln + gweeks +
#                                   pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f + moud.yes.f,
#                                 data=basic, fitter = cph, xtrans= cph_multi_i, n.impute=20, y=TRUE, x=TRUE, surv=TRUE)

cph_multi_spm <- cph(Surv(days.to.event, outcome.bin) ~ rcs(delivery_year,3) + total.rx.days.exp3 +  momage  + routedeliv.new + gweeks +  
                   pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE) 

```

```{r}
# print(summary(robcov(cph_multi_sp, cluster = basic$unique_id.x), delivery_year=c(2016, 2020),  momage=c(25, 30), rpl_themes=c(0.5, 0.75), total.rx.days.exp=c(0,1), vnames = 'labels'), digits=3)
# print(anova(robcov(cph_multi_sp), vnames="labels"), digits=3)
# x <- Predict(cph_multi_sp, delivery_year=c(2007:2020))
# x$yhat_exp <- exp(x$yhat)
# x$lower_exp <- exp(x$lower)
# x$upper_exp <- exp(x$upper)

print(summary(robcov(cph_multi_spm, cluster = moud$unique_id.x), momage=c(27,31),  total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39,37),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2007,2015),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2010,2014),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2015, 2017),  vnames = 'labels'))
# 
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2015, 2020),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2017, 2020),  vnames = 'labels'))
# print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2017, 2018),  vnames = 'labels'))

print(anova(robcov(cph_multi_spm), vnames="labels"), digits=3)

```

```{r,fig.height=6, fig.width=8}
plot(summary(cph_multi_spm, momage=c(25, 35),   total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), vnames = 'labels'))
```

```{r, fig.height=12, fig.width=12}
ggplot(Predict(cph_multi_spm, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical")

```

<br>

## Linear model (only buprenorphine)
```{r, include=FALSE, warning=FALSE, message=FALSE}
# cph_multi_sp <- fit.mult.impute(Surv(follow.up.days.log10, death.or.od) ~  rcs(delivery_year, 3)  + total.rx.days.exp  + momage + momraceeth_comb.new + rpl_themes + routedeliv.new + previouspregnanciestotaln + gweeks +
#                                   pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f + moud.yes.f,
#                                 data=basic, fitter = cph, xtrans= cph_multi_i, n.impute=20, y=TRUE, x=TRUE, surv=TRUE)

onlybup <- moud[!is.na(moud$total.rx.days.total_3bup), ]
cph_multi_spm2 <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.total_3bup +  momage  + routedeliv.new + gweeks +  
                   pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE) 

```

```{r}

print(summary(robcov(cph_multi_spm2, cluster = moud$unique_id.x), momage=c(27,31),  total.rx.days.total_3bup=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39, 37),  vnames = 'labels'))

print(anova(robcov(cph_multi_spm2), vnames="labels"), digits=3)

```

<br>

### Univariate 
```{r}
cph_uni2 <- cph(Surv(days.to.event, outcome.bin) ~ total.rx.days.total_3bup , data=moud, x=TRUE, y=TRUE) 
print(summary(robcov(cph_uni2, cluster = moud$unique_id.x), total.rx.days.total_3bup=c(1, 31),  vnames = 'labels'))

print(anova(robcov(cph_uni2), vnames="labels"), digits=3)


```


# By MOUD timing {-}
# Survival analysis
## KM Curves
```{r}
kmout_moud <- survfit(Surv(days.to.event, outcome.bin) ~moud.both3div, data=moud)
# test with survminer
# survminer::ggsurvplot(kmout, data=basic, fun="event")

#KM survival curve for only 1 year
plot(kmout_moud, conf.int=FALSE, col=c("red", "black", "blue"), xlab="Days Since 41-days Post-Partum", ylab="Probability of Death or Overdose", main= "KM curves by MOUD timing", xlim=c(0, 323), fun="event")
legend('bottomright', c("Pre-delivery", "Post-delivery", "Both"), col=c("red","black", "blue"), lty=1)


kable(data.frame(kmout_moud$time, kmout_moud$surv), col.names = c("time", "probability"))
```


## Log-Rank Test

```{r}
surv_diff_moud <- survdiff(Surv(days.to.event, outcome.bin) ~moud.both3div, data=moud)
print(surv_diff_moud)
```
<br>

## Multivariable Cox PH model
## Linear continuous terms (timing + days) {-}
```{r, warning=FALSE, message=NA, include=FALSE}

dd <- datadist(moud)
options(datadist= 'dd')
dd$limits$moud.both3div[2] <- "Both pre and post"

cph_multim_time <- cph(Surv(days.to.event, outcome.bin) ~ moud.both3div + total.rx.days.exp3 + delivery_year +  momage +  routedeliv.new + gweeks +  
                    pain.combine.f +  alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE) 

```

```{r}

print(summary(robcov(cph_multim_time, cluster = moud$unique_id.x), momage=c(27,31), total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39, 37),
              vnames = 'labels'))
print(anova(robcov(cph_multim_time), vnames="labels"), digits=2)

```

```{r, fig.height=6, fig.width=8}
plot(summary(cph_multim, momage=c(27, 31), total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39, 37),  vnames = 'labels'))
```

```{r, fig.height=12, fig.width=12}
ggplot(Predict(cph_multim, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical")
```

<br>

## Timing only {-}
```{r}

cph_multim_time2 <- cph(Surv(days.to.event, outcome.bin) ~ moud.both3div + delivery_year +  momage +  routedeliv.new + gweeks +  
                    pain.combine.f +  alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE)


```


```{r}

print(summary(robcov(cph_multim_time2, cluster = moud$unique_id.x), momage=c(27,31),  delivery_year=c(2016, 2018), gweeks=c(39, 37),
              vnames = 'labels'))
print(anova(robcov(cph_multim_time2), vnames="labels"), digits=2)

```

<br>

## Linear continuous terms (timing + days BUP) {-}
```{r, warning=FALSE, message=NA, include=FALSE}

cph_multim_time_b <- cph(Surv(days.to.event, outcome.bin) ~ moud.both3div + total.rx.days.total_3bup + delivery_year +  momage +  routedeliv.new + gweeks +  
                    pain.combine.f +  alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE) 

```

```{r}

print(summary(robcov(cph_multim_time_b, cluster = moud$unique_id.x), momage=c(25,35), total.rx.days.total_3bup=c(1, 31), delivery_year=c(2014, 2018), 
              vnames = 'labels'))
print(anova(robcov(cph_multim_time_b), vnames="labels"), digits=2)

```

<br>

## Timing only BUP {-}
```{r}
moudbup <- moud[!is.na(moud$total.rx.days.total_3bup),]
cph_multim_time_b2 <- cph(Surv(days.to.event, outcome.bin) ~ moud.both3div + delivery_year +  momage +  routedeliv.new + gweeks +  
                    pain.combine.f +  alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moudbup, x=TRUE, y=TRUE)


```


```{r}

print(summary(robcov(cph_multim_time_b2, cluster = moudbup$unique_id.x), momage=c(25,35),  delivery_year=c(2014, 2018), 
              vnames = 'labels'))
print(anova(robcov(cph_multim_time_b2), vnames="labels"), digits=2)

```

<br>

# Sensitivity analysis 2007-2018
## Descriptive Statistics

* There are 5728 pregnancies in 5049 unique individuals
```{r}
moud18 <- moud[which(moud$delivery_year <= 2018), ]

tangram(paste('1~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] +
          final.outcome.f + final.outcome2.f + total.rx.days.exp[0] + total.rx.days.b90[0] + total.rx.days.p41[0] + 
          total.rx.days.exp2[0] + total.rx.days.b90_2[0] + total.rx.days.p41_2[0] + 
          total.rx.days.exp3[0] + total.rx.days.b90_3[0] + total.rx.days.p41_3[0] + avg.dose.pre[2] + avg.dose.post[2] + avg.dose.total[2] +
          total.rx.days.b90_3bup[2] + total.rx.days.p41_3bup[2] +  total.rx.days.total_3bup[2] + avg.dose.pre.bup[2] + avg.dose.post.bup[2] + avg.dose.total.bup[2] +
          pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +',
              noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'), 
        data=moud18,  pformat=4, digits=2, style="nejm")
```


### Time to event for overdose and death
```{r}
deathod18 <- moud18[which(moud18$final.outcome.f %in% c("Death", "Overdose")), ]
describe(deathod18$days.to.event)

death18 <- deathod18[which(deathod18$final.outcome.f=="Death"), ]
describe(death18$days.to.event)

od18 <- deathod18[which(deathod18$final.outcome.f=="Overdose"), ]
describe(od18$days.to.event)
```

<br>


## Survival Analysis {-}
## Kaplan Meier Plot 
```{r}
kmoutm18 <- survfit(Surv(days.to.event, outcome.bin) ~1, data=moud18)
print(kmoutm18)
# test with survminer
# survminer::ggsurvplot(kmout, data=basic, fun="event")

#KM survival curve for only 1 year
# plot(kmoutm, conf.int=FALSE, col=c("red", "blue"), xlab="Days since 41-days post-partum", ylab="Probability of Death or Overdose", main= "KM curves by MOUD timing", xlim=c(0, 365), fun="event")
# legend('bottomright', c("Either pre or post", "Both pre and post"), col=c("red", "blue"), lty=1)

plot(kmoutm18, conf.int=FALSE, col=c("red", "blue"), xlab="Days since 41-days post-partum", ylab="Probability of Death or Overdose", xlim=c(0, 323), fun="event")

summary(kmoutm18$time)

kable(data.frame(kmoutm18$time, kmoutm18$surv), col.names = c("time", "probability"))
```


## Multivariable Cox PH model {-}
### Linear continuous terms
```{r, warning=FALSE, message=NA, include=FALSE}

dd <- datadist(moud18)
options(datadist= 'dd')
# dd$limits$pain.combine.f[2] <- "No"
# dd$limits$alcuse.othersub.f[2] <- "No"
# dd$limits$moud.both3[2] <- "Either pre or post"
# 
cph_multim18 <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.exp3 +  momage +  routedeliv.new + gweeks +  
                    pain.combine.f +  alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud18, x=TRUE, y=TRUE) 
# cox.zph(cph_multim) # PH assumption met
```

```{r}

print(summary(robcov(cph_multim18, cluster = moud18$unique_id.x), momage=c(27,31), total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39,37),
              vnames = 'labels'))
print(anova(robcov(cph_multim18), vnames="labels"), digits=2)

```


```{r, fig.height=6, fig.width=8}
plot(summary(cph_multim18, momage=c(25, 35),   total.rx.days.exp3=c(1, 31), delivery_year=c(2016, 2018),  vnames = 'labels'))
```

`

```{r, fig.height=12, fig.width=12}
ggplot(Predict(cph_multim18, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical")
```


<br>
  
### Univariable HR
```{r}
cph_uni18 <- cph(Surv(days.to.event, outcome.bin) ~ total.rx.days.exp3 , data=moud18, x=TRUE, y=TRUE) 

print(summary(robcov(cph_uni18, cluster = moud18$unique_id.x),  total.rx.days.exp3=c(1, 31), vnames = 'labels'))
print(anova(robcov(cph_uni18), vnames="labels"), digits=2)
```


# Sensitivity analysis shift-forward
```{r}
tangram(1~  total.rx.days.exp3[0] + total.rx.days.shifted[0], 
        data=moud, pformat=4, digits=2, style="nejm")
```

## Survival Analysis 
## Multivariable Cox PH model {-}
### Linear continuous terms {-}
```{r, warning=FALSE, message=NA, include=FALSE}
dd <- datadist(moud)
options(datadist= 'dd')

cph_multis <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.shifted +  momage +  routedeliv.new + gweeks +  
                    pain.combine.f +  alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE) 
# cox.zph(cph_multim) # PH assumption met
```

```{r}

print(summary(robcov(cph_multis, cluster = moud$unique_id.x), momage=c(27,31), total.rx.days.shifted=c(1, 31), delivery_year=c(2016, 2018), gweeks=c(39,37),
              vnames = 'labels'))
print(anova(robcov(cph_multis), vnames="labels"), digits=2)

```


### Univariable HR
```{r}
cph_unis <- cph(Surv(days.to.event, outcome.bin) ~  total.rx.days.shifted , data=moud, x=TRUE, y=TRUE) 

print(summary(robcov(cph_unis, cluster = moud$unique_id.x),   total.rx.days.shifted =c(1, 31), vnames = 'labels'))
print(anova(robcov(cph_unis), vnames="labels"), digits=2)
```


<!-- ## With splines (only buprenorphine) -->
<!-- ## Excluding few variables -->
<!-- ```{r, include=FALSE, warning=FALSE, message=FALSE} -->

<!-- cph_multi_spm3 <- cph(Surv(days.to.event, outcome.bin) ~ rcs(delivery_year,3) + total.rx.days.total_3bup +  momage  +  -->
<!--                    pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE)  -->

<!-- ``` -->

<!-- ```{r} -->

<!-- print(summary(robcov(cph_multi_spm3, cluster = moud$unique_id.x), momage=c(25,35),  total.rx.days.total_3bup=c(1, 31), delivery_year=c(2014, 2019),  vnames = 'labels')) -->

<!-- print(anova(robcov(cph_multi_spm3), vnames="labels"), digits=3) -->

<!-- ``` -->

<!-- ## With splines (on outcome without all payer data) -->
<!-- ```{r, include=FALSE, warning=FALSE, message=FALSE} -->
<!-- moud$outcome.bin2 <- ifelse(moud$outcome2 %in% c("Death", "Overdose"), 1, 0) -->
<!-- cph_multi_sp2 <- cph(Surv(days.to.event, outcome.bin2) ~ rcs(delivery_year,3) + total.rx.days.exp3 +  momage + routedeliv.new + gweeks +   -->
<!--   pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE)  -->

<!-- ``` -->

<!-- ```{r} -->
<!-- print(summary(robcov(cph_multi_sp2, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014, 2019),  vnames = 'labels')) -->
<!-- print(anova(robcov(cph_multi_sp2), vnames="labels"), digits=3) -->

<!-- ``` -->

<!-- # Propensity Score (Matching Weights) by MOUD timing -->

<!-- ```{r} -->
<!-- moud$moud.both3.num <- as.numeric(moud$moud.both3) -->
<!-- moud$moud.both3.num <- moud$moud.both3.num - 1 # has to be 0 or 1 -->

<!-- form.ps2 <- "moud.both3.num ~ delivery_year +  momage + routedeliv.new.num + gweeks + pain.combine  + alcuse.othersub + benzo.anticonv + anx.dep.comb + mat_morbidity" -->

<!-- V.name2 <- c("delivery_year", "momage" , "routedeliv.new.num" , "gweeks", "pain.combine", "alcuse.othersub", "benzo.anticonv", "anx.dep.comb", "mat_morbidity") -->

<!-- mw2 <- psw(data=moud, form.ps = form.ps2, weight="MW", mirror.hist = TRUE, V.name = V.name2, add.weight = TRUE)   -->

<!-- ps2balance <- psw.balance(data= moud, weight="MW", form=form.ps2, V.name = V.name2) -->

<!-- ps2balancebefore <- as.data.frame(ps2balance$std.diff.before) -->
<!-- ps2balancebefore$std.diff.pct.prop <- round(abs(ps2balancebefore$std.diff.pct)/100, digits=2) -->
<!-- ps2balanceafter <- as.data.frame(ps2balance$std.diff.after) -->
<!-- ps2balanceafter$std.diff.pct.prop <- round(abs(ps2balanceafter$std.diff.pct)/100, digits=2) -->
<!-- ps2balance.to.plot <- data.frame(var= rownames(ps2balancebefore), Before= ps2balancebefore$std.diff.pct.prop, After= ps2balanceafter$std.diff.pct.prop) -->
<!-- ps2balance.to.plot$Variable <- c("Delivery year", "Age", "Route of delivery", "Gestational age", "Pain condition combined", "Alc/substance use", "Benzo/anticonvulsants", "Anxiety/Depression", "Maternal morbidity") -->

<!-- ps2balance.to.plot <- melt(ps2balance.to.plot, id="Variable") -->
<!-- ps2balance.to.plot <- subset(ps2balance.to.plot, variable!= "var") -->
<!-- ps2balance.to.plot$value <- as.numeric(ps2balance.to.plot$value) -->
<!-- ps2balance.to.plot$Variable <- with(ps2balance.to.plot, reorder(x= Variable, X= value, FUN= sum)) -->
<!-- ``` -->
<!-- <br> -->


<!-- * Mirror histograms show propensity scores of the control group above the horizontal line in white, while the treatment group is below the horizontal line in dark green. Light green bars display the summation of weights of subjects within each propensity score stratum. This plot shows good overlap between the control and treatment groups, and the close similarity between them after applying matching weights.  -->



<!-- ## Absolute standardized mean difference plot -->
<!-- ```{r} -->
<!-- ggplot(ps2balance.to.plot, aes(x=value, y=Variable), fill=variable) + -->
<!--   geom_point(aes(color= variable), size=3) + labs(x= "SMD", y="") + theme_minimal() + theme(legend.title = element_blank()) -->

<!-- ``` -->
<!-- <br> -->

<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog2 <- glm(outcome.bin ~ moud.both3, family=binomial(link="logit"), weights = mw2$W, data=moud) -->
<!-- summary(pslog2) -->
<!-- ``` -->

<!-- <br> -->

<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog2), confint(pslog2))) -->
<!-- ``` -->
<!-- <br> -->


<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog2moud <- glm(outcome.bin ~ total.rx.days.exp3, family=binomial(link="logit"), weights = mw2$W, data=moud) -->
<!-- summary(pslog2moud) -->
<!-- ``` -->
<!-- <br> -->


<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog2moud), confint(pslog2moud))) -->
<!-- ``` -->
<!-- <br> -->

<!-- * OR for 30-day supply: 0.99902160^30 equals: -->
<!-- ```{r} -->
<!-- 0.99902160^30 -->
<!-- ``` -->

<!-- <br> -->



<!-- # Propensity score excluding naltrexone/methadone by MOUD receipt (2015+) -->

<!-- ```{r} -->
<!-- # variables need to be numeric for PSW later -->
<!-- moud2015 <- moud[moud$delivery_year>=2015, ] -->
<!-- moud2015 <- moud[!(moud$last_rx_type %in% c("Methadone", "Naltrexone", "Naltrexone Hydrochloride", "Naltrexone, Naltrexone Hydrochloride")), ] -->
<!-- moud2015$last_rx_type <- droplevels(moud2015$last_rx_type) -->


<!-- form.ps2.2015 <- "moud.both3.num ~ delivery_year +  momage + routedeliv.new.num + gweeks + pain.combine  + alcuse.othersub + benzo.anticonv + anx.dep.comb + mat_morbidity" -->

<!-- V.name2.2015 <- c("delivery_year", "momage" , "routedeliv.new.num" , "gweeks", "pain.combine", "alcuse.othersub", "benzo.anticonv", "anx.dep.comb", "mat_morbidity") -->

<!-- mw2.2015 <- psw(data=moud2015, form.ps = form.ps2.2015, weight="MW", mirror.hist = TRUE, V.name = V.name2.2015, add.weight = TRUE)   -->

<!-- ps2.2015balance <- psw.balance(data= moud2015, weight="MW", form=form.ps2.2015, V.name = V.name2.2015) -->

<!-- ps2.2015balancebefore <- as.data.frame(ps2.2015balance$std.diff.before) -->
<!-- ps2.2015balancebefore$std.diff.pct.prop <- round(abs(ps2.2015balancebefore$std.diff.pct)/100, digits=2.2015) -->
<!-- ps2.2015balanceafter <- as.data.frame(ps2.2015balance$std.diff.after) -->
<!-- ps2.2015balanceafter$std.diff.pct.prop <- round(abs(ps2.2015balanceafter$std.diff.pct)/100, digits=2.2015) -->
<!-- ps2.2015balance.to.plot <- data.frame(var= rownames(ps2.2015balancebefore), Before= ps2.2015balancebefore$std.diff.pct.prop, After= ps2.2015balanceafter$std.diff.pct.prop) -->
<!-- ps2.2015balance.to.plot$Variable <- c("Delivery year", "Age", "Route of delivery", "Gestational age", "Pain condition combined", "Alc/substance use", "Benzo/anticonvulsants", "Anxiety/Depression", "Maternal morbidity") -->

<!-- ps2.2015balance.to.plot <- melt(ps2.2015balance.to.plot, id="Variable") -->
<!-- ps2.2015balance.to.plot <- subset(ps2.2015balance.to.plot, variable!= "var") -->
<!-- ps2.2015balance.to.plot$value <- as.numeric(ps2.2015balance.to.plot$value) -->
<!-- ps2.2015balance.to.plot$Variable <- with(ps2.2015balance.to.plot, reorder(x= Variable, X= value, FUN= sum)) -->
<!-- ``` -->



<!-- * Mirror histograms show propensity scores of the control group above the horizontal line in white, while the treatment group is below the horizontal line in dark green. Light green bars display the summation of weights of subjects within each propensity score stratum. This plot shows good overlap between the control and treatment groups, and the close similarity between them after applying matching weights.  -->



<!-- ## Absolute standardized mean difference plot -->
<!-- ```{r} -->
<!-- ggplot(ps2.2015balance.to.plot, aes(x=value, y=Variable), fill=variable) + -->
<!--   geom_point(aes(color= variable), size=3) + labs(x= "SMD", y="") + theme_minimal() + theme(legend.title = element_blank()) -->

<!-- ``` -->


<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog2.2015 <- glm(outcome.bin ~ moud.both3.num, family=binomial(link="logit"), weights = mw2.2015$W, data=moud2015) -->

<!-- summary(pslog2.2015) -->
<!-- ``` -->
<!-- * Exposure is period (both versus only one period relative to delivery) -->

<!-- <br> -->


<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog2.2015), confint(pslog2.2015))) -->
<!-- ``` -->

<!-- <br> -->

<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog2moud2015 <- glm(outcome.bin ~ total.rx.days.exp3, family=binomial(link="logit"), weights = mw2.2015$W, data=moud2015) -->
<!-- summary(pslog2moud2015) -->
<!-- ``` -->
<!-- <br> -->


<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog2moud2015), confint(pslog2moud2015))) -->
<!-- ``` -->
<!-- <br> -->

<!-- * OR for 30-day supply: 0.99890050^30 equals: -->
<!-- ```{r} -->
<!-- 0.99890050^30 -->
<!-- ``` -->


<!-- <br> -->

<!-- ```{r} -->
<!-- moud$moud.group.lastrx <- ifelse(moud$last_rx_type %in% c("Buprenorphine", "Buprenorphine Hydrochloride",  -->
<!--                                                           "Buprenorphine, Buprenorphine Hydrochloride"), "Buprenorphine", -->
<!--                                  ifelse(moud$last_rx_type %in% c("Buprenorphine Hydrochloride, Buprenorphine/Naloxone", "Buprenorphine/Naloxone"), "Buprenorphine/Naloxone", NA)) -->
<!-- moud$moud.group.lastrx.num <- ifelse(moud$moud.group.lastrx=="Buprenorphine", 0, 1) -->
<!-- ``` -->

<!-- <br> -->


<!-- # Propensity Score (Matching Weights) by last MOUD type -->

<!-- ```{r} -->
<!-- form.ps3 <- "moud.group.lastrx.num ~ delivery_year +  momage + routedeliv.new.num + gweeks + pain.combine  + alcuse.othersub + benzo.anticonv + anx.dep.comb + mat_morbidity" -->

<!-- V.name3 <- c("delivery_year", "momage" , "routedeliv.new.num" , "gweeks", "pain.combine", "alcuse.othersub", "benzo.anticonv", "anx.dep.comb", "mat_morbidity") -->

<!-- moud2 <- moud[!is.na(moud$moud.group.lastrx), ] -->

<!-- mw3 <- psw(data=moud2, form.ps = form.ps3, weight="MW", mirror.hist = TRUE, V.name = V.name3, add.weight = TRUE)   -->

<!-- ps3balance <- psw.balance(data= moud, weight="MW", form=form.ps3, V.name = V.name3) -->

<!-- ps3balancebefore <- as.data.frame(ps3balance$std.diff.before) -->
<!-- ps3balancebefore$std.diff.pct.prop <- round(abs(ps3balancebefore$std.diff.pct)/100, digits=2) -->
<!-- ps3balanceafter <- as.data.frame(ps3balance$std.diff.after) -->
<!-- ps3balanceafter$std.diff.pct.prop <- round(abs(ps3balanceafter$std.diff.pct)/100, digits=2) -->
<!-- ps3balance.to.plot <- data.frame(var= rownames(ps3balancebefore), Before= ps3balancebefore$std.diff.pct.prop, After= ps3balanceafter$std.diff.pct.prop) -->
<!-- ps3balance.to.plot$Variable <- c("Delivery year", "Age", "Route of delivery", "Gestational age", "Pain condition combined", "Alc/substance use", "Benzo/anticonvulsants", "Anxiety/Depression", "Maternal morbidity") -->

<!-- ps3balance.to.plot <- melt(ps3balance.to.plot, id="Variable") -->
<!-- ps3balance.to.plot <- subset(ps3balance.to.plot, variable!= "var") -->
<!-- ps3balance.to.plot$value <- as.numeric(ps3balance.to.plot$value) -->
<!-- ps3balance.to.plot$Variable <- with(ps3balance.to.plot, reorder(x= Variable, X= value, FUN= sum)) -->
<!-- ``` -->

<!-- <br> -->


<!-- * Mirror histograms show propensity scores of the control group above the horizontal line in white, while the treatment group is below the horizontal line in dark green. Light green bars display the summation of weights of subjects within each propensity score stratum. This plot shows good overlap between the control and treatment groups, and the close similarity between them after applying matching weights.  -->



<!-- ## Absolute standardized mean difference plot -->
<!-- ```{r} -->
<!-- ggplot(ps3balance.to.plot, aes(x=value, y=Variable), fill=variable) + -->
<!--   geom_point(aes(color= variable), size=3) + labs(x= "SMD", y="") + theme_minimal() + theme(legend.title = element_blank()) -->

<!-- ``` -->

<!-- <br> -->

<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog3 <- glm(outcome.bin ~ moud.group.lastrx.num, family=binomial(link="logit"), weights = mw3$W, data=moud2) -->
<!-- summary(pslog3) -->
<!-- ``` -->
<!-- * Reference = Buprenorphine only -->
<!-- * Comparison group = Buprenorphine/Naloxone -->

<!-- <br> -->

<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog3), confint(pslog3))) -->
<!-- ``` -->
<!-- <br> -->


<!-- # Propensity Score (Matching Weights) by last MOUD type (2015+) -->

<!-- ```{r} -->
<!-- form.ps3 <- "moud.group.lastrx.num ~ delivery_year +  momage + routedeliv.new.num + gweeks + pain.combine  + alcuse.othersub + benzo.anticonv + anx.dep.comb + mat_morbidity" -->

<!-- V.name3 <- c("delivery_year", "momage" , "routedeliv.new.num" , "gweeks", "pain.combine", "alcuse.othersub", "benzo.anticonv", "anx.dep.comb", "mat_morbidity") -->
<!-- moud2015 <- moud2[moud2$delivery_year>=2015, ] -->
<!-- moud2015$last_rx_type <- droplevels(moud2015$last_rx_type) -->

<!-- mw3.2015 <- psw(data=moud2015, form.ps = form.ps3, weight="MW", mirror.hist = TRUE, V.name = V.name3, add.weight = TRUE)   -->

<!-- ps3balance2015 <- psw.balance(data= moud2015, weight="MW", form=form.ps3, V.name = V.name3) -->

<!-- ps3balancebefore2015 <- as.data.frame(ps3balance2015$std.diff.before) -->
<!-- ps3balancebefore2015$std.diff.pct.prop <- round(abs(ps3balancebefore2015$std.diff.pct)/100, digits=2) -->
<!-- ps3balanceafter2015 <- as.data.frame(ps3balance2015$std.diff.after) -->
<!-- ps3balanceafter2015$std.diff.pct.prop <- round(abs(ps3balanceafter2015$std.diff.pct)/100, digits=2) -->
<!-- ps3balance.to.plot2015 <- data.frame(var= rownames(ps3balancebefore2015), Before= ps3balancebefore2015$std.diff.pct.prop, After= ps3balanceafter$std.diff.pct.prop) -->
<!-- ps3balance.to.plot2015$Variable <- c("Delivery year", "Age", "Route of delivery", "Gestational age", "Pain condition combined", "Alc/substance use", "Benzo/anticonvulsants", "Anxiety/Depression", "Maternal morbidity") -->

<!-- ps3balance.to.plot2015 <- melt(ps3balance.to.plot2015, id="Variable") -->
<!-- ps3balance.to.plot2015 <- subset(ps3balance.to.plot2015, variable!= "var") -->
<!-- ps3balance.to.plot2015$value <- as.numeric(ps3balance.to.plot2015$value) -->
<!-- ps3balance.to.plot2015$Variable <- with(ps3balance.to.plot2015, reorder(x= Variable, X= value, FUN= sum)) -->
<!-- ``` -->


<!-- <br> -->


<!--   * Mirror histograms show propensity scores of the control group above the horizontal line in white, while the treatment group is below the horizontal line in dark green. Light green bars display the summation of weights of subjects within each propensity score stratum. This plot shows good overlap between the control and treatment groups, and the close similarity between them after applying matching weights.  -->



<!-- ## Absolute standardized mean difference plot -->
<!-- ```{r} -->
<!-- ggplot(ps3balance.to.plot2015, aes(x=value, y=Variable), fill=variable) + -->
<!--   geom_point(aes(color= variable), size=3) + labs(x= "SMD", y="") + theme_minimal() + theme(legend.title = element_blank()) -->

<!-- ``` -->


<!-- <br> -->

<!-- ## Logistic regression model -->
<!-- ```{r, warning=FALSE} -->
<!-- pslog3.2015 <- glm(outcome.bin ~ moud.group.lastrx.num, family=binomial(link="logit"), weights = mw3.2015$W, data=moud2015) -->
<!-- summary(pslog3.2015) -->
<!-- ``` -->

<!-- `` -->
<!-- * Reference = Buprenorphine only -->
<!-- * Comparison group = Buprenorphine/Naloxone -->

<!-- <br> -->

<!-- * exponentiated Odds Ratio -->
<!-- ```{r, warning=FALSE} -->
<!-- exp(cbind(OR = coef(pslog3.2015), confint(pslog3.2015))) -->
<!-- ``` -->
<!-- <br> -->


<!-- <br> -->

<!-- # Subgroup analysis  -->
<!-- ## Both periods only -->
<!-- ```{r} -->
<!-- moud_both <- moud[moud$moud.both3=="Both pre and post", ] -->
<!-- cph_multi_sp_both <- cph(Surv(days.to.event, outcome.bin) ~ rcs(delivery_year,3) + total.rx.days.exp3 +  momage + momraceeth_comb.new + routedeliv.new + gweeks +   -->
<!--                    pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f + rpl_themes + -->
<!--                    previouspregnanciestotaln, data=moud_both, x=TRUE, y=TRUE)  -->


<!-- print(summary(robcov(cph_multi_sp_both, cluster = moud_both$unique_id.x), momage=c(25, 35),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014, 2019), rpl_themes=c(0.25, 0.75),  vnames = 'labels')) -->
<!-- print(anova(robcov(cph_multi_sp_both), vnames="labels"), digits=2) -->
<!-- ``` -->

<!-- <br> -->


<!-- ## One period only -->

<!-- ```{r} -->
<!-- moud_one <- moud[moud$moud.both3=="Either pre or post", ] -->
<!-- cph_multi_sp_one <- cph(Surv(days.to.event, outcome.bin) ~ rcs(delivery_year,3) + total.rx.days.exp3 +  momage + momraceeth_comb.new + routedeliv.new + gweeks +   -->
<!--                    pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f + rpl_themes + -->
<!--                    previouspregnanciestotaln, data=moud_one, x=TRUE, y=TRUE)  -->


<!-- print(summary(robcov(cph_multi_sp_one, cluster = moud_one$unique_id.x),  momage=c(25, 35),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014, 2019), rpl_themes=c(0.25, 0.75),  vnames = 'labels')) -->
<!-- print(anova(robcov(cph_multi_sp_one), vnames="labels"), digits=2) -->
<!-- ``` -->

<!-- <br> -->

<!-- ```{r} -->
<!-- extract180 <- af_extract[af_extract$ndc %in% c(ndc$ndcnum, 'Q9991',    -->
<!-- 'Q9992',   -->
<!-- 'J0571', -->
<!-- 'J0572', -->
<!-- 'J0573', -->
<!-- 'J0574', -->
<!-- 'J0575', -->
<!-- 'G2068', -->
<!-- 'G2069', -->
<!-- 'G2070', -->
<!-- 'G2072', -->
<!-- 'G2079'), ] -->
<!-- # test <- table(af_extract$ndc) -->
<!-- # tail(test, 20) -->

<!-- extract180 <- merge(extract180, ndc, by.x="ndc", by.y= "ndcnum") -->
<!-- extract180 <- extract180[extract180$BTrack==180, ] -->

<!-- extract <- extract180 -->
<!-- # subset people who are in the exposure group -->
<!-- # basic <- af_master_new[af_master_new$StudyID %in% extract$StudyID, ] -->
<!-- basic <- af_master[af_master$Expose1==1, ] -->

<!-- # want 180 now -->
<!-- basic <- basic[basic$BTrack==180, ] -->
<!-- basic$filledrx <- ifelse(basic$StudyID %in% extract$StudyID, 1, 0) -->

<!-- af_master$unique_id <- sub(".*-", "", af_master$StudyID) -->
<!-- basic$unique_id <- sub(".*-", "", basic$StudyID) -->

<!-- # merge covariates -->
<!-- af_cov$unique_id <- sub(".*-", "", af_cov$StudyID) -->
<!-- # only include one of the pregnancies if twins/triplets  -->
<!-- basic <- basic[order(basic$unique_id, basic$StudyID), ] -->
<!-- # separate those with multiple pregnancies -->
<!-- # create indicators: if next row covariates are the same as previous row then they are twins/triplets -->
<!-- basic <- basic %>% mutate(rownum= row_number()) %>% data.frame() -->
<!-- # possibly twins?  -->
<!-- # these are the records where the study id is duplicated (either twins or multiple children over years) -->
<!-- twins <- basic %>% group_by(unique_id) %>%  -->
<!--             mutate(n = n()) %>% # count number of same study id -->
<!--                 filter(n> 1) %>% -->
<!--                         mutate(mom_dob_prevrow = lag(MomDOB_day),  -->
<!--                             gweeks20_prevrow = lag(GWeeks20_dt_day)) -->
<!-- twins$multbirths <- ifelse(twins$mom_dob_prevrow == twins$MomDOB_day & twins$gweeks20_prevrow == twins$GWeeks20_dt_day, 1, 0) -->
<!-- twins$multbirths <- ifelse(is.na(twins$multbirths), 0, twins$multbirths)   -->
<!-- not_twins <- twins[twins$multbirths==0 ,] -->
<!-- twins2 <- twins[twins$multbirths==1,] # these are twins/triplets -->

<!-- # exclude twins/triplets -->
<!-- basic <- basic[! (basic$rownum %in% twins2$rownum),  ] -->


<!-- # some covariates are in Btrack 180 only and not 90 -->
<!-- af_cov <- as.data.frame(af_cov) -->
<!-- # af_cov2 <- af_cov[af_cov$BTrack==90, ] -->
<!-- af_cov2 <- af_cov[af_cov$StudyID %in% basic$StudyID, ] -->

<!-- af_cov_u <- af_cov2[ , !(colnames(af_cov2) %in% c("BTrack", "condition", "condition_from_day", "CodeID"))] -->
<!-- af_cov_u <- af_cov_u[!duplicated(af_cov_u), ] -->

<!-- basic <- merge(basic, af_cov_u, by= "StudyID", all.x = TRUE) -->

<!-- # df <- basic #creating a copy to work off of just in case -->
<!-- names(basic) <- tolower(names(basic)) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- # duplicated ids  -->
<!-- # dups <- basic[which(duplicated(basic$unique_id)), "unique_id"] -->
<!-- # dup <- basic[which(duplicated(basic$unique_id)), ] -->
<!-- # dupdf <- basic[which(basic$unique_id %in% dups), ] -->
<!-- # include only the first pregnancy for now  -->

<!-- # Format covariates -->
<!-- label(basic$momage) <- "Age at delivery" -->
<!-- basic$momrace <- ifelse(basic$momrace=="", NA, basic$momrace) -->
<!-- basic$momrace_desc <- ifelse(basic$momrace_desc=="", NA, basic$momrace_desc) -->
<!-- basic$momethnicity_desc <- ifelse(basic$momethnicity_desc=="", NA, basic$momethnicity_desc) -->
<!-- basic$momrace.new <- ifelse(basic$momrace_desc %in% c("White", "Bridged White"), "White", -->
<!--                             ifelse(basic$momrace_desc %in% c("Black or African-American", "Bridged Black"), "Black of African American", -->
<!--                                    ifelse(basic$momrace_desc %in% c("American Indian or Alaskan Native", "Bridged American Indian or Alaskan Native"), "American Indian or Alaska Native", -->
<!--                                       ifelse(basic$momrace_desc %in% c("Asian Indian", "Bridged Asian or Pacific Isander", "Chinese", "Filipino", "Korean", "Native Hawaiian", "Other Asian", "Vietnamese"), "Asian or Pacific Islander", -->
<!--                                              ifelse(basic$momrace_desc %in% c("Other Race", "Refused or Unknown"), "Other or Unknown", NA))))) -->

<!-- label(basic$momrace.new) <- "Race" -->

<!-- basic$momhisp.new <- ifelse(basic$momethnicity_desc %in% c("\tNot Hispanic", "\tOther", "\tUnknown"), "Not Hispanic",  -->
<!--                             ifelse(is.na(basic$momethnicity_desc), NA, "Hispanic")) -->

<!-- basic$momraceeth.new <- ifelse(basic$momhisp.new=="Hispanic", "Hispanic", -->
<!--                                ifelse((basic$momhisp.new!="Hispanic"| is.na(basic$momhisp.new)) & basic$momrace.new=="Asian or Pacific Islander", "Non-Hispanic Asian", -->
<!--                                       ifelse((basic$momhisp.new!="Hispanic"| is.na(basic$momhisp.new)) & basic$momrace.new=="Black of African American", "Non-Hispanic Black", -->
<!--                                              ifelse((basic$momhisp.new!="Hispanic"| is.na(basic$momhisp.new)) & basic$momrace.new=="White", "Non-Hispanic White", NA)))) -->
<!-- basic$momraceeth.new <- ifelse(basic$momhisp.new=="Not Hispanic" & is.na(basic$momrace.new) & is.na(basic$momraceeth.new), "Non-Hispanic Other",  -->
<!--                                ifelse(is.na(basic$momhisp.new) & basic$momrace.new=="White" & is.na(basic$momraceeth.new), "Non-Hispanic White", -->
<!--                                       ifelse(is.na(basic$momhisp.new) & basic$momrace.new %in% c("American Indian or Alaska Native", "Other or Unknown"), "Non-Hispanic Other",  -->
<!--                                              ifelse(basic$momhisp.new=="Not Hispanic" & basic$momrace.new %in% c("American Indian or Alaska Native", "Other or Unknown") & is.na(basic$momraceeth.new), "Non-Hispanic Other", basic$momraceeth.new)))) -->
<!-- label(basic$momraceeth.new) <- "Race/Ethnicity combined" -->

<!-- label(basic$rpl_themes) <- "CDC SVI overall percentile ranking of all 4 themes" -->


<!-- #  information related to each delivery/pregnancy -->
<!-- basic$routedeliv.new <- ifelse(basic$deliverycesarean=="Y", "Cesarean", ifelse(basic$deliveryvaginal=="Y", "Vaginal", ifelse(basic$deliveryforcep=="Y", "Vaginal", ifelse(basic$deliveryvacuum=="Y", "Vaginal", NA)))) -->
<!-- label(basic$routedeliv.new) <- "Route of delivery" -->

<!-- # exclude if route of delivery everything is marked as "NO" -->
<!-- # excluded one person -->
<!-- basic <- basic[!is.na(basic$routedeliv.new), ] -->

<!-- label(basic$previouspregnanciestotaln) <- "Number of previous pregnancies" -->
<!-- label(basic$cat4_seg1) <- "Enrollment up to 365 days post-partum" -->

<!-- basic$enroll365.new.f <- factor(basic$cat4_seg1, levels = c(0,1), labels=c("No", "Yes")) -->
<!-- label(basic$enroll365.new.f) <- "Enrollment up to 365 days post-partum" -->

<!-- label(basic$momhisp.new) <- "Hispanic" -->

<!-- # Death -->
<!-- # want to capture deaths from day 42 and onwards but before loss of enrollment; some deaths captured after they disenroll (for now, but will go back and capture them later) -->
<!-- basic$death.new <- ifelse(!is.na(basic$momdeath_day) & basic$momdeath_day <= 323 & basic$momdeath_day <= basic$fdate_seg1_day, 1, 0) -->
<!-- basic$death.new.f <- factor(basic$death.new, levels=c(0, 1), labels=c("No", "Yes")) -->
<!-- label(basic$death.new.f) <- "Death during follow-up" -->
<!-- ``` -->


<!-- ```{r} -->
<!-- # comorbidities -->
<!-- af_cov2 <- af_cov2[af_cov2$condition!="", ] -->
<!-- af_cov2 <- af_cov2[af_cov2$BTrack==180, ] -->

<!-- # transform conditions from long to wide -->
<!-- af_cov3 <- af_cov2[ , c("StudyID", "condition_from_day", "condition")] -->

<!-- comorb <- af_cov3 %>% group_by(StudyID) %>% -->
<!--   pivot_wider(id_cols = "StudyID", names_from = "condition", values_from = "condition") %>% -->
<!--   mutate_all(~replace(., . == "NULL", 0)) %>% -->
<!--   mutate_all(~replace(., . != "0", 1)) %>%  -->
<!--   mutate_if(is.list, as.numeric) %>% -->
<!--   ungroup() %>% as.data.frame() -->

<!-- basic <- merge(basic, comorb, by.x = "studyid", by.y = "StudyID", all.x = TRUE) -->

<!-- comorbidities <- names(comorb)[2:17] -->
<!-- comorblabs <- gsub("_", " ", comorbidities) -->
<!-- comorblabs <- capitalize(comorblabs) -->
<!-- comorblabs <- gsub("nsaids", "NSAIDS", comorblabs) -->
<!-- comorblabs <- gsub("substance", "substance use", comorblabs) -->
<!-- comorblabs <- gsub("Bipolard", "Bipolar disorder", comorblabs) -->
<!-- comorblabs <- gsub("psych", "psychotic disorders", comorblabs) -->
<!-- comorblabs <- gsub("benzo", "benzodiazepine use", comorblabs) -->
<!-- comorblabs <- gsub("usedis", "use disorder", comorblabs) -->
<!-- comorblabs <- gsub("anticonv", "anticonvulsant use", comorblabs) -->
<!-- comorblabs <- gsub("Personality dis", "Personality disorder", comorblabs) -->
<!-- comorblabs <- gsub("Alc", "Alcohol", comorblabs) -->
<!-- comorblabs <- gsub("Mat", "Maternal", comorblabs) -->
<!-- comorblabs <- gsub("relax", "relaxant use", comorblabs) -->


<!-- # create factor variables -->
<!-- for(i in 1:length(comorbidities)) { -->
<!--   basic[[paste0(comorbidities[i], ".f")]] <- basic[[comorbidities[i]]] -->
<!--   basic[[paste0(comorbidities[i], ".f")]] <- ifelse(is.na(basic[[paste0(comorbidities[i], ".f")]]), 0, basic[[paste0(comorbidities[i], ".f")]]) -->
<!--   basic[[paste0(comorbidities[i], ".f")]] <- factor(basic[[paste0(comorbidities[i], ".f")]], levels = c(0:1), labels=c('No', 'Yes')) -->
<!--   label(basic[[paste0(comorbidities[i], ".f")]]) <- comorblabs[i] -->
<!-- } -->

<!-- label(basic$gweeks) <- "Number of Gestation weeks" -->

<!-- # Combine variables for analyses -->
<!-- basic$pain.combine <- ifelse(basic$back_pain.f=="Yes" | basic$dental_pain.f=="Yes" | basic$fibromyalgia.f=="Yes" | basic$med_nsaids.f=="Yes" | basic$med_muscle_relax.f=="Yes", 1, 0) -->
<!-- basic$pain.combine.f <- factor(basic$pain.combine, levels = 0:1, labels=c("No", "Yes")) -->
<!-- label(basic$pain.combine.f) <- "Combined pain" -->

<!-- basic$personality.combine <- ifelse(basic$personality_dis.f=="Yes"| basic$schizophrenia_psych.f=="Yes", 1, 0) -->
<!-- basic$personality.combine.f <- factor(basic$personality.combine, levels = 0:1, labels=c("No", "Yes")) -->
<!-- label(basic$personality.combine.f) <- "Schizophrenia or personality disorder" -->

<!-- basic$alcuse.othersub <- ifelse(basic$alc_usedis.f=="Yes" | basic$other_substance.f=="Yes", 1, 0) -->
<!-- basic$alcuse.othersub.f <- factor(basic$alcuse.othersub, levels = 0:1, labels=c("No", "Yes")) -->
<!-- label(basic$alcuse.othersub.f) <- "Alcohol or other substance use" -->

<!-- basic$benzo.anticonv <- ifelse(basic$med_benzo.f=="Yes" | basic$med_anticonv.f == "Yes", 1, 0) -->
<!-- basic$benzo.anticonv.f <- factor(basic$benzo.anticonv, levels=0:1, labels=c("No", "Yes")) -->
<!-- label(basic$benzo.anticonv.f) <- "Benzodiazepine or anticonvulsant use" -->

<!-- ``` -->

<!-- ```{r} -->
<!-- # overdose -->
<!-- af_out <- af_outcome[af_outcome$BTrack==180 & af_outcome$StudyID %in% basic$studyid, ] -->
<!-- names(af_out) <- tolower(names(af_out)) -->
<!-- # change to count outcomes that occur after disenrollment too -->
<!-- af_out$fdate_seg1_day <- basic[match(af_out$studyid, basic$studyid), "fdate_seg1_day"] -->
<!-- # check to see which overdose or death occurs after disenrollment -->
<!-- # x <- af_out[which(af_out$overdose_day - 42 > af_out$fdate_seg1_day), ] -->
<!-- # subset events that occured before disenrollment (before) -->
<!-- af_out$momdeath_day <- basic[match(af_out$studyid, basic$studyid), "momdeath_day"] -->
<!-- # af_out <- af_out[(!is.na(af_out$overdose_day) & (af_out$overdose_day <= af_out$fdate_seg1_day)) | (!is.na(af_out$momdeath_day) & (af_out$momdeath_day<= af_out$fdate_seg1_day)), ] -->

<!-- # outcomes file contains multiple overdose records (previous ODs can show up on next visits) -->
<!-- # summarize information on total number of overdoses -->
<!-- af_out2 <- af_out %>% group_by(studyid) %>% -->
<!--   filter(!is.na(overdose_day)) %>% -->
<!--   mutate(n= n()) %>% data.frame() -->

<!-- basic$num.od <- af_out2[match(basic$studyid, af_out2$studyid), "n"] -->
<!-- label(basic$num.od) <- "Number of any overdoses recorded during follow-up" -->

<!-- # Filter out opioid related overdose only -->
<!-- af_out_op <- af_out %>% group_by(studyid) %>% -->
<!--   filter(overdose_opioid==1) %>% -->
<!--   mutate(n= n()) %>% data.frame() -->

<!-- basic$num.opioid.od <- af_out_op[match(basic$studyid, af_out_op$studyid), "n"] -->
<!-- label(basic$num.opioid.od) <- "Number of opioid overdoses recorded during follow-up"   -->

<!-- # ever had od during follow-up -->
<!-- basic$outcome.od <- ifelse(!is.na(basic$num.od), 1, 0) -->
<!-- label(basic$outcome.od) <- "Had any overdose during follow-up" -->

<!-- # ever had opioid od during follow-up -->
<!-- basic$outcome.opioid.od <- ifelse(!is.na(basic$num.opioid.od), 1, 0) -->
<!-- label(basic$outcome.opioid.od) <- "Had any opioid overdose during follow-up" -->


<!-- # More deaths from outcome table; merge into large table -->
<!-- af_out_death <- af_out %>% group_by(studyid) %>%  -->
<!--   filter(!is.na(momdeath_day)) %>% data.frame() -->

<!-- af_out_death <- af_out_death[ , c("studyid", grep("dcause", names(af_out_death), value=TRUE), "momdeath_day")] -->
<!-- af_out_death <- unique(af_out_death) -->
<!-- # basic <- merge(basic, af_out_death, by="studyid", all.x = TRUE) -->
<!-- basic$death_day <- basic$momdeath_day -->
<!-- basic$death_day <- ifelse(basic$death_day > 323, NA, basic$death_day) -->

<!-- # see dead info only -->
<!-- #dead <- basic[which(!is.na(basic$death_day) | basic$death.new==1| !is.na(basic$momdeath_day)), c("studyid", grep("death", names(basic), value=TRUE), grep("dcause", names(test), value=TRUE))] -->
<!-- #  -->
<!-- # # create composite outcome of death or overdose only -->
<!-- # basic$death.or.od <- ifelse(!is.na(basic$num.od) | basic$death.new==1, 1, 0) -->
<!-- # basic$death.or.op.od <-  ifelse(!is.na(basic$num.opioid.od) | basic$death.new==1, 1, 0) -->
<!-- #  -->
<!-- # basic$death.or.od.f <- factor(basic$death.or.od, levels=0:1, labels = c("No", "Yes")) -->
<!-- # label(basic$death.or.od.f) <- "Death or any overdose during follow-up" -->
<!-- #  -->
<!-- # basic$death.or.op.od <- factor(basic$death.or.op.od, levels=0:1, labels = c("No", "Yes")) -->
<!-- # label(basic$death.or.op.od) <- "Death or any opioid overdose during follow-up" -->
<!-- # #  -->
<!-- # ignore below for now -->
<!-- #  -->
<!-- # basic$death.cause.new <- substr(basic$mrecip_dcause, 1,3) -->
<!-- # basic$death.cause.new <- ifelse(basic$death.cause.new=="", NA, basic$death.cause.new) -->
<!-- #  -->
<!-- # names(deathcode)[1] <- "Code" -->
<!-- # names(deathcode)[2] <- "Value" -->
<!-- #  -->
<!-- # basic$death.cause.new2 <- deathcode[match(basic$death.cause.new, deathcode$Code), "Value"] -->
<!-- # basic$death.cause.new2[basic$death.new==0 & !is.na(basic$death.cause.new2)] <- NA -->
<!-- #  -->
<!-- # # Opioid-related and non-opioid related overdose deaths have T40 ICD codes -->
<!-- # # none in this data -->
<!-- # # table(basic$death.cause.new) -->
<!-- # basic$death.od.fu.new <- 0 -->
<!-- # basic$death.od.fu.new <- factor(basic$death.od.fu.new, levels=0:1, labels=c("No", "Yes")) -->
<!-- # label(basic$death.od.fu.new) <- "Death from overdose during follow-up" -->

<!-- ``` -->


<!-- ```{r} -->
<!-- # get days of OUD medications -->
<!-- # 180 days before delivery -->
<!-- extract180 <- extract[extract$Rx_B180_DOB==1 & extract$Expose1==1, ] -->
<!-- names(extract180) <- tolower(names(extract180)) -->
<!-- extract180 <- extract180[extract180$studyid %in% basic$studyid, ] -->
<!-- extract180 <- extract180[order(extract180$studyid, extract180$from), ] -->


<!-- # summarize the number of days supplied cap at 180+42 days = 132 days (Method 1) -->
<!-- extract180 <- extract180 %>% group_by(studyid) %>%  -->
<!--                   mutate(total.rx.days = cumsum(daysup)) %>% data.frame() -->

<!-- #extract180[which(extract180$studyid=="0020018032034-04544465-03494573"), c("studyid", "daysup", "from", "total.rx.days")] -->
<!-- extract180$total.rx.days <- ifelse(extract180$total.rx.days > 180, 180, extract180$total.rx.days) -->

<!-- extract180s <- extract180 %>% group_by(studyid) %>% slice_max(total.rx.days, with_ties = FALSE) %>% data.frame() -->


<!-- # 41 days after delivery -->
<!-- extract41 <- extract[extract$Rx_DOB_P41==1 & extract$Expose1==1, ] -->
<!-- names(extract41) <- tolower(names(extract41)) -->
<!-- extract41 <- extract41[extract41$studyid %in% basic$studyid, ] -->
<!-- extract41 <- extract41[order(extract41$studyid, extract41$from), ] -->

<!-- extract41 <- extract41 %>% group_by(studyid) %>%  -->
<!--                   mutate(total.rx.days = cumsum(daysup)) %>% data.frame() -->
<!-- extract41$total.rx.days <- ifelse(extract41$total.rx.days > 42, 42, extract41$total.rx.days) -->

<!-- extract41s <- extract41 %>% group_by(studyid) %>% slice_max(total.rx.days, with_ties = FALSE) %>% data.frame() -->


<!-- basic$total.rx.days.b180 <- extract180s[match(basic$studyid, extract180s$studyid), "total.rx.days"] -->
<!-- basic$total.rx.days.p41 <- extract41s[match(basic$studyid, extract41s$studyid), "total.rx.days"] -->

<!-- basic$total.rx.days.b180 <- ifelse(is.na(basic$total.rx.days.b180), 0, basic$total.rx.days.b180) -->
<!-- basic$total.rx.days.b180 <- ifelse(basic$total.rx.days.b180 > 180, 180, basic$total.rx.days.b180) -->
<!-- basic$total.rx.days.p41 <- ifelse(is.na(basic$total.rx.days.p41), 0, basic$total.rx.days.p41) -->



<!-- # Method 2 -->
<!-- # Add up total days supplied, if > 180 then add that number to days after delivery instead of capping at 180 -->
<!-- extract180_2 <- extract180 %>% group_by(studyid) %>%  -->
<!--                   mutate(total.rx.days = cumsum(daysup), -->
<!--                          max.rx.days = max(total.rx.days)) %>% data.frame() -->

<!-- extract180_2$total.rx.minus180 <- ifelse(extract180_2$max.rx.days > 180, extract180_2$max.rx.days - 180, 0) -->


<!-- extract41_2 <- extract41 %>% group_by(studyid) %>%  -->
<!--                   mutate(total.rx.days = cumsum(daysup), -->
<!--                          max.rx.days = max(total.rx.days)) %>% data.frame() -->

<!-- basic$total.rx.days.b180_2 <- extract180_2[match(basic$studyid, extract180_2$studyid), "max.rx.days"] -->
<!-- basic$total.rx.days.b180_2 <- ifelse(basic$total.rx.days.b180_2 > 180, 180, basic$total.rx.days.b180_2) -->
<!-- basic$total.rx.days.b180_2 <- ifelse(is.na(basic$total.rx.days.b180_2), 0, basic$total.rx.days.b180_2) -->
<!-- basic$total.rx.days.b180_minus180 <- extract180_2[match(basic$studyid, extract180_2$studyid), "total.rx.minus180"] -->
<!-- basic$total.rx.days.b180_minus180 <- ifelse(is.na(basic$total.rx.days.b180_minus180), 0, basic$total.rx.days.b180_minus180) -->

<!-- basic$total.rx.days.p41_2max <- extract41_2[match(basic$studyid, extract41_2$studyid), "max.rx.days"] -->
<!-- basic$total.rx.days.p41_2max <- ifelse(basic$total.rx.days.p41_2max > 42, 42, basic$total.rx.days.p41_2max) -->
<!-- basic$total.rx.days.p41_2max <- ifelse(is.na(basic$total.rx.days.p41_2max), 0, basic$total.rx.days.p41_2max) -->

<!-- basic$total.rx.days.p41_2 <- rowSums(basic[, c("total.rx.days.p41_2max", "total.rx.days.b180_minus180")]) -->
<!-- basic$total.rx.days.p41_2 <- ifelse(basic$total.rx.days.p41_2 > 42, 42, basic$total.rx.days.p41_2) -->

<!-- # Method 3 -->
<!-- # counting for overlapping prescription dates -->
<!-- extract180_3 <- extract180 %>% group_by(studyid) %>% -->
<!--     arrange(studyid, from_day) %>% -->
<!--     mutate(indx = c(0, cumsum(as.numeric(lead(from_day)) > -->
<!--                      cummax(as.numeric(from_day + daysup)))[-n()])) %>% -->
<!--     group_by(studyid, indx) %>% -->
<!--     summarise(start = min(from_day), end = max(from_day + daysup)) -->

<!-- extract180_3 <- extract180_3 %>% group_by(studyid, indx) %>% -->
<!--     mutate(sum_days = abs(start - end))  -->

<!-- extract180_3 <- extract180_3 %>% group_by(studyid) %>% -->
<!--     mutate(total_days = sum(sum_days)) -->

<!-- extract180_3 <- as.data.frame(extract180_3) -->

<!-- extract41_3 <- extract41 %>% group_by(studyid) %>% -->
<!--     arrange(studyid, from_day) %>% -->
<!--     mutate(indx = c(0, cumsum(as.numeric(lead(from_day)) > -->
<!--                      cummax(as.numeric(from_day + daysup)))[-n()])) %>% -->
<!--     group_by(studyid, indx) %>% -->
<!--     summarise(start = min(from_day), end = max(from_day + daysup)) -->

<!-- # have to cap it to -1 otherwise going into after exposure period -->
<!-- extract41_3$end <- ifelse(extract41_3$end > -1, -1, extract41_3$end) -->

<!-- extract41_3 <- extract41_3 %>% group_by(studyid, indx) %>% -->
<!--     mutate(sum_days = abs(start) - abs(end) +1 )  -->

<!-- extract41_3 <- extract41_3 %>% group_by(studyid) %>% -->
<!--     mutate(total_days = sum(sum_days)) -->

<!-- extract41_3 <- as.data.frame(extract41_3) -->



<!-- basic$total.rx.days.b180_3 <- extract180_3[match(basic$studyid, extract180_3$studyid), "total_days"] -->
<!-- basic$total.rx.days.p41_3 <- extract41_3[match(basic$studyid, extract41_3$studyid), "total_days"] -->
<!-- basic$total.rx.days.b180_3 <- ifelse(is.na(basic$total.rx.days.b180_3), 0, basic$total.rx.days.b180_3) -->
<!-- basic$total.rx.days.p41_3 <- ifelse(is.na(basic$total.rx.days.p41_3), 0, basic$total.rx.days.p41_3) -->

<!-- # if greater than 180 then add to after delivery days supplied -->
<!-- basic$total.rx.days.p41_3 <- ifelse(basic$total.rx.days.b180_3 > 180, (basic$total.rx.days.b180_3 - 180) + basic$total.rx.days.p41_3, basic$total.rx.days.p41_3) -->
<!-- basic$total.rx.days.b180_3 <- ifelse(basic$total.rx.days.b180_3 > 180, 180, basic$total.rx.days.b180_3) -->
<!-- basic$total.rx.days.p41_3 <- ifelse(basic$total.rx.days.b180_3 > 42, 42, basic$total.rx.days.p41_3) -->


<!-- basic$total.rx.days.exp <- rowSums(basic[ , c("total.rx.days.b180", "total.rx.days.p41")]) -->
<!-- basic$total.rx.days.exp2 <- rowSums(basic[ , c("total.rx.days.b180_2", "total.rx.days.p41_2")]) -->
<!-- basic$total.rx.days.exp3 <- rowSums(basic[ , c("total.rx.days.b180_3", "total.rx.days.p41_3")]) -->

<!-- label(basic$total.rx.days.b180) <- "Total days supplied MOUD before delivery" -->
<!-- label(basic$total.rx.days.p41) <- "Total days supplied MOUD after delivery" -->
<!-- label(basic$total.rx.days.exp) <- "Total days supplied MOUD" -->

<!-- # summarize very last medication received per person -->
<!-- extract_last <- extract %>% group_by(StudyID) %>% slice_max(from_day) %>% data.frame -->
<!-- names(extract_last) <- tolower(names(extract_last)) -->
<!-- ## summarize the type of medication and number of prescriptions  -->
<!-- extract_last <- extract_last[ , c("studyid", "from_day", "gennme")] -->
<!-- extract_last <- unique(extract_last) -->
<!-- extract_last <- reshape2::dcast(extract_last, studyid ~ gennme) -->
<!-- extract_last$last_rx_type <- apply(extract_last[, 2:6], 1, function(x) toString(na.omit(x))) -->
<!-- names(extract_last) <- make.names(names(extract_last)) -->

<!-- ## summarize the type of medication and number of prescriptions  -->
<!-- extract_u <- extract -->
<!-- names(extract_u) <- tolower(names(extract_u)) -->
<!-- extract_u <- extract_u[ , c("studyid", "from_day", "gennme")] -->
<!-- extract_u <- unique(extract_u) -->
<!-- extract_u <- reshape2::dcast(extract_u, studyid ~ gennme) -->
<!-- names(extract_u) <- make.names(names(extract_u)) -->
<!-- extract_u <- extract_u %>%  mutate_if(is.numeric, ~1 * (. > 0)) -->
<!-- extract_u$unique.rx <- rowSums(extract_u[ , 2:6], na.rm = TRUE) -->

<!-- basic$last_rx_type <- extract_last[match(basic$studyid, extract_last$studyid), "last_rx_type"] -->
<!-- basic$unique.rx <- extract_u[match(basic$studyid, extract_u$studyid), "unique.rx"] -->

<!-- basic$last_rx_type <- as.factor(basic$last_rx_type) -->
<!-- label(basic$last_rx_type) <- "Last medication type" -->

<!-- label(basic$unique.rx) <- "Unique number of prescriptions" -->

<!-- basic$moud.yes <- ifelse(basic$studyid %in% extract$StudyID, 1, 0) -->
<!-- basic$moud.yes.f <- factor(basic$moud.yes, levels=0:1, labels=c("No MOUD", "MOUD")) -->
<!-- label(basic$moud.yes.f) <- "MOUD" -->

<!-- ``` -->

<!-- ```{r} -->
<!-- # follow-up time if they had an event (counting events not captured in claims like all hospital payer data) -->
<!-- # if death or overdose happens after 365 days, then use overdose day or last observed day -->
<!-- af_out365 <- af_out[af_out$momdeath_day <= 323 | af_out$overdose_day <= 323 | af_out$fdate_seg1_day <= 323, ] -->
<!-- af_out365 <- af_out365 %>% group_by(studyid) %>% slice_min(overdose_day, with_ties = FALSE) %>% data.frame() -->

<!-- af_out365$days.to.event <- pmin(af_out365$momdeath_day, af_out365$overdose_day, na.rm = TRUE) -->
<!-- af_out365$outcome <- ifelse(af_out365$days.to.event== af_out365$momdeath_day , "Death",  -->
<!--                             ifelse(af_out365$days.to.event == af_out365$overdose_day, "Overdose", NA)) -->
<!-- af_out365$outcome <- ifelse(is.na(af_out365$outcome) & is.na(af_out365$momdeath_day) & !is.na(af_out365$overdose_day), "Overdose", af_out365$outcome) -->

<!-- af_out365 <- as.data.frame(af_out365) -->
<!-- af_out <- as.data.frame(af_out) -->
<!-- af_out$days.to.event <- af_out365[match(af_out$studyid, af_out365$studyid), "days.to.event"] -->
<!-- basic$days.to.event <- af_out[match(basic$studyid, af_out365$studyid), "days.to.event"] -->
<!-- basic$outcome <- af_out365[match(basic$studyid, af_out365$studyid), "outcome"] -->
<!-- basic$days.to.event <- ifelse(is.na(basic$days.to.event), basic$fdate_seg1_day, basic$days.to.event) -->
<!-- basic$days.to.event <- ifelse(basic$days.to.event > 365, 365, basic$days.to.event) -->
<!-- label(basic$days.to.event) <- "Days to event (w/payer data)" -->

<!-- af_out$days.to.event2 <- af_out365[match(af_out$studyid, af_out365$studyid), "days.to.event2"] -->
<!-- basic$days.to.event2 <- af_out[match(basic$studyid, af_out365$studyid), "days.to.event2"] -->
<!-- basic$outcome <- af_out365[match(basic$studyid, af_out365$studyid), "outcome"] -->
<!-- basic$days.to.event2 <- ifelse(is.na(basic$days.to.event2), basic$fdate_seg1_day, basic$days.to.event2) -->
<!-- basic$days.to.event2 <- ifelse(basic$days.to.event2 > 365, 365, basic$days.to.event2) -->
<!-- label(basic$days.to.event2) <- "Days to event2 (w/o all payer data)" -->




<!-- # Loss of enrollment within 1 year of delivery -->
<!-- basic$loss.enroll <- ifelse(basic$cat1_seg1==1 & basic$fdate_seg1_day +42 <=365, 1, 0) -->
<!-- label(basic$loss.enroll) <- "Loss of enrollment within 1-year of delivery" -->

<!-- basic$outcome.num <- ifelse(is.na(basic$outcome), 0, 1) -->

<!-- basic$final.outcome <- ifelse(basic$outcome =="Death", 1, -->
<!--                               ifelse(basic$outcome=="Overdose", 2,  -->
<!--                                      ifelse(basic$loss.enroll==1 & is.na(basic$outcome), 3, NA))) -->
<!-- basic$final.outcome <- ifelse(is.na(basic$final.outcome) & basic$loss.enroll==1, 3, basic$final.outcome) -->
<!-- basic$final.outcome <- ifelse(is.na(basic$final.outcome) & basic$loss.enroll!=1, 4, basic$final.outcome) -->



<!-- # check: basic[which(basic$death.new==1), c("studyid", "days.to.event", "follow.up.days", "loss.enroll",  "death_day", "fdate_seg1_day", "death.or.od", "death.new", "final.outcome") ] -->
<!-- # some people have fdate that are greater than death_day (ex: ID "002007049737-02925139-05210809") as well as those with fdate before death date (maybe death recorded later on after disenrollment?) -->

<!-- basic$final.outcome.f <- factor(basic$final.outcome, levels=1:4, labels=c("Death", "Overdose", "Loss of enrollment", "Censored")) -->
<!-- label(basic$final.outcome.f) <- "Final outcome" -->


<!-- ``` -->


<!-- ```{r} -->
<!-- # follow-up times calculated without accounting for OD after disenrollment  -->
<!-- af_out365$days.to.event2 <- pmin(af_out365$overdose_day, af_out365$fdate_seg1_day, af_out365$momdeath_day) -->
<!-- af_out365$days.to.event2 <- ifelse(is.na(af_out365$overdose_day), pmin(af_out365$fdate_seg1_day, af_out365$momdeath_day), af_out365$days.to.event2) -->
<!-- af_out365$days.to.event2 <- ifelse(is.na(af_out365$momdeath_day), pmin(af_out365$fdate_seg1_day, af_out365$overdose_day), af_out365$days.to.event2) -->

<!-- af_out365$outcome2 <- ifelse(af_out365$days.to.event2 == af_out365$momdeath_day, "Death", -->
<!--                              ifelse(af_out365$days.to.event2 == af_out365$overdose_day, "Overdose", -->
<!--                                     ifelse(af_out365$days.to.event2 == af_out365$fdate_seg1_day, "Loss of enrollment", "Censored"))) -->
<!-- af_out365$outcome2 <- ifelse(is.na(af_out365$outcome2) & af_out365$days.to.event == af_out365$days.to.event2, as.character(af_out365$outcome), af_out365$outcome2) -->
<!-- af_out365$outcome2 <- ifelse(is.na(af_out365$outcome2) & af_out365$days.to.event2 == af_out365$fdate_seg1_day, "Loss of enrollment", af_out365$outcome2) -->

<!-- basic$outcome2 <- af_out365[match(basic$studyid, af_out365$studyid), "outcome2"] -->

<!-- basic$final.outcome2 <- ifelse(basic$outcome2 =="Death", 1, -->
<!--                               ifelse(basic$outcome2=="Overdose", 2,  -->
<!--                                      ifelse(basic$loss.enroll==1 & is.na(basic$outcome2), 3, NA))) -->
<!-- basic$final.outcome2 <- ifelse(is.na(basic$final.outcome2) & basic$loss.enroll==1, 3, basic$final.outcome2) -->
<!-- basic$final.outcome2 <- ifelse(is.na(basic$final.outcome2) & basic$loss.enroll!=1, 4, basic$final.outcome2) -->

<!-- basic$final.outcome2.f <- factor(basic$final.outcome2, levels=1:4, labels=c("Death", "Overdose", "Loss of enrollment", "Censored")) -->
<!-- label(basic$final.outcome2.f) <- "Final outcome (w/o all payer data)" -->

<!-- ``` -->



<!-- ```{r} -->
<!-- # Additional covariates to be adjusted for -->
<!-- basic$delivery_year <- as.integer(basic$delivery_year) -->
<!-- label(basic$delivery_year) <- "Year of delivery" -->

<!-- basic$delivery.year.cat <- as.factor(basic$delivery_year) -->
<!-- label(basic$delivery.year.cat) <- "Year of delivery as categorical" -->

<!-- basic$momraceeth_comb.new <- as.character(basic$momraceeth.new) -->
<!-- basic$momraceeth_comb.new <- ifelse(basic$momraceeth_comb.new %nin% c("Non-Hispanic White"), "All Other", basic$momraceeth_comb.new) -->
<!-- basic$momraceeth_comb.new <- as.factor(basic$momraceeth_comb.new) -->
<!-- label(basic$momraceeth_comb.new) <- "Race/Ethnicity binary" -->

<!-- label(basic$total.rx.days.b180_2) <- "Total days supplied MOUD before delivery (method 2)" -->
<!-- label(basic$total.rx.days.p41_2) <- "Total days supplied MOUD after delivery (method 2)" -->

<!-- label(basic$total.rx.days.b180_3) <- "Total days supplied MOUD before delivery (method 3)" -->
<!-- label(basic$total.rx.days.p41_3) <- "Total days supplied MOUD after delivery (method 3)" -->

<!-- basic$moud.pre <- ifelse(basic$total.rx.days.b180 >0, 1, 0) -->
<!-- basic$moud.pre2 <- ifelse(basic$total.rx.days.b180_2 >0, 1, 0) -->
<!-- basic$moud.pre3 <- ifelse(basic$total.rx.days.b180_3 >0, 1, 0) -->

<!-- basic$moud.post <- ifelse(basic$total.rx.days.p41 >0, 1, 0) -->
<!-- basic$moud.post2 <- ifelse(basic$total.rx.days.p41_2 >0, 1, 0) -->
<!-- basic$moud.post3 <- ifelse(basic$total.rx.days.p41_3 >0, 1, 0) -->

<!-- basic$moud.both <- ifelse(basic$moud.pre==1 & basic$moud.post==1, 1, 0) -->
<!-- basic$moud.both2 <- ifelse(basic$moud.pre2==1 & basic$moud.post2==1, 1, 0) -->
<!-- basic$moud.both3 <- ifelse(basic$moud.pre3==1 & basic$moud.post3==1, 1, 0) -->
<!-- basic$moud.onlypre3 <- ifelse(basic$moud.pre3==1 & basic$moud.post3==0, 1, 0) -->
<!-- basic$moud.onlypost3 <- ifelse(basic$moud.post3==1 & basic$moud.pre3==0, 1, 0) -->

<!-- basic$moud.both3div <- ifelse(basic$moud.onlypre3==1, 1,  -->
<!--                               ifelse(basic$moud.onlypost3==1, 2,  -->
<!--                                      ifelse(basic$moud.both3==1, 3, 0))) -->
<!-- basic$moud.both3div <- factor(basic$moud.both3div, levels=0:3, labels=c("No MOUD", "Only pre", "Only post", "Both pre and post")) -->
<!-- label(basic$moud.both3div) <- "MOUD timing (method 3)" -->

<!-- basic$moud.pre <- factor(basic$moud.pre, levels=0:1, labels=c("No MOUD", "MOUD")) -->
<!-- label(basic$moud.pre) <- "MOUD pre-delivery (method 1)" -->

<!-- basic$moud.pre2 <- factor(basic$moud.pre2, levels=0:1, labels=c("No MOUD", "MOUD")) -->
<!-- label(basic$moud.pre2) <- "MOUD pre-delivery (method 2)" -->

<!-- basic$moud.pre3 <- factor(basic$moud.pre3, levels=0:1, labels=c("No MOUD", "MOUD")) -->
<!-- label(basic$moud.pre3) <- "MOUD pre-delivery (method 3)" -->

<!-- basic$moud.post <- factor(basic$moud.post, levels=0:1, labels=c("No MOUD", "MOUD")) -->
<!-- label(basic$moud.post) <- "MOUD post-delivery (method 1)" -->

<!-- basic$moud.post2 <- factor(basic$moud.post2, levels=0:1, labels=c("No MOUD", "MOUD")) -->
<!-- label(basic$moud.post2) <- "MOUD post-delivery (method 2)" -->

<!-- basic$moud.post3 <- factor(basic$moud.post3, levels=0:1, labels=c("No MOUD", "MOUD")) -->
<!-- label(basic$moud.post3) <- "MOUD post-delivery (method 3)" -->

<!-- basic$moud.both <- factor(basic$moud.both, levels=0:1, labels=c("Either pre or post", "Both pre and post")) -->
<!-- label(basic$moud.both) <- "MOUD pre and post (method 1)" -->

<!-- basic$moud.both2 <- factor(basic$moud.both2, levels=0:1, labels=c("Either pre or post", "Both pre and post")) -->
<!-- label(basic$moud.both2) <- "MOUD pre and post (method 2)" -->

<!-- basic$moud.both3 <- factor(basic$moud.both3, levels=0:1, labels=c("Either pre or post", "Both pre and post")) -->
<!-- label(basic$moud.both3) <- "MOUD pre and post (method 3)" -->




<!-- label(basic$total.rx.days.exp2) <- "Total days supplied MOUD (method 2)" -->
<!-- label(basic$total.rx.days.exp3) <- "Total days supplied MOUD (method 3)" -->

<!-- basic$anx.dep.comb <- ifelse(basic$anxiety==1 | basic$depression== 1, 1, 0) -->
<!-- basic$anx.dep.comb <- ifelse(is.na(basic$anx.dep.comb), 0, basic$anx.dep.comb) -->
<!-- basic$anx.dep.comb.f <- factor(basic$anx.dep.comb, levels=0:1, labels=c("No", "Yes")) -->
<!-- label(basic$anx.dep.comb.f) <- "Anxiety and depression combined" -->

<!-- basic$outcome.bin <- ifelse(basic$final.outcome %in% c(1, 2), 1, 0) -->

<!-- moud <- basic[basic$moud.yes==1, ] -->
<!-- ``` -->



<!-- # Descriptive Statistics (180 days pre-delivery & up to 5 days loss enroll) -->
<!-- ## Demographics/Clinical Characteristics -->
<!-- ### Column percents -->

<!-- <br> -->

<!-- * There are 4247 pregnancies in 3837 unique individuals.  -->
<!-- ```{r} -->
<!-- tangram(paste('1~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] + -->
<!--           final.outcome.f + final.outcome2.f + total.rx.days.exp[0] + total.rx.days.b180[0] + total.rx.days.p41[0] +  -->
<!--           total.rx.days.exp2[0] + total.rx.days.b180_2[0] + total.rx.days.p41_2[0] +  -->
<!--           total.rx.days.exp3[0] + total.rx.days.b180_3[0] + total.rx.days.p41_3[0] + -->
<!--           pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f +', -->
<!--             noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'),  -->
<!--         data=moud,  pformat=4, digits=2, style="nejm") -->
<!-- ``` -->

<!-- <br> -->

<!-- ### MOUD information -->

<!-- ```{r} -->
<!-- tangram(1~ moud.pre + moud.pre2 + moud.pre3 + moud.post + moud.post2 + moud.post3 + -->
<!--           moud.both + moud.both2 + moud.both3, data= moud, pformat=4, digits=2, style= "nejm", collapse_single=FALSE) -->
<!-- ``` -->
<!-- <br> -->



<!-- ### Characteristics based on MOUD timing (method 3) -->
<!-- ```{r} -->
<!-- tangram(paste('moud.both3~ delivery_year[0] + delivery.year.cat + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] + -->
<!--           final.outcome.f + total.rx.days.exp[0] + total.rx.days.b180[0] + total.rx.days.p41[0] +  -->
<!--           total.rx.days.exp2[0] + total.rx.days.b180_2[0] + total.rx.days.p41_2[0] +  -->
<!--           total.rx.days.exp3[0] + total.rx.days.b180_3[0] + total.rx.days.p41_3[0] + -->
<!--           pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f + anx.dep.comb.f + last_rx_type+', -->
<!--             noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'),  -->
<!--         data=moud, test=TRUE, pformat=4, digits=2, style="nejm") -->
<!-- ``` -->

<!-- <br> -->



<!-- <!-- ### Row Percents --> -->
<!-- <!-- ```{r} --> -->
<!-- <!-- tangram(paste('moud.yes.f~ delivery_year[0] + momage[0] +gweeks[0] + momrace.new + momhisp.new + momraceeth.new + routedeliv.new  + previouspregnanciestotaln[0] + --> -->
<!-- <!--           death.new.f + death.or.od.f + death.or.op.od + final.outcome.f + total.rx.days.exp[0] + total.rx.days.b180[0] + total.rx.days.p41[0] +  --> -->
<!-- <!--           pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + mat_morbidity.f +', --> -->
<!-- <!--             noquote(paste0(comorbidities, ".f", collapse="+")), '+ rpl_themes'),  --> -->
<!-- <!--         data=basic,  overall=TRUE, test=TRUE, pformat=4, digits=2, style="nejm", row_percents=TRUE) --> -->
<!-- <!-- ``` --> -->

<!-- <br> -->

<!-- ## Histogram of MOUD days supplied -->
<!-- ### Before delivery up to 180 days (Method 1) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.b180)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 180, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- ### Before delivery up to 180 days (Method 2) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.b180_2)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 180, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- <br> -->

<!-- ### Before delivery up to 180 days (Method 3) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.b180_3)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 180, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- <br> -->

<!-- ### After delivery up to 41 days (Method 1) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.p41_2)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 40, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- ### After delivery up to 41 days (Method 3) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.p41_2)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 40, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- <br> -->

<!-- ### After delivery up to 41 days (Method 3) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.p41_3)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 40, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->


<!-- <br> -->

<!-- ### Entire exposure (Method 1) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.exp)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 230, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- <br> -->


<!-- ### Entire exposure (Method 2) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.exp2)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 230, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- <br> -->

<!-- ### Entire exposure (Method 3) -->
<!-- ```{r} -->
<!-- ggplot(moud, aes(x=total.rx.days.exp3)) +  -->
<!--   geom_histogram(color="black", fill="white") + -->
<!--   scale_x_continuous(breaks = seq(0, 230, 10)) +  -->
<!--   labs(x= "Days", y="Count") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- <br> -->

<!-- ```{r, fig.width=10, fig.height= 5} -->
<!-- custom_x <- list( -->
<!--   scale_x_continuous(limits=c(0, 180)), -->
<!--   scale_x_continuous(limits=c(0, 42)), -->
<!--   scale_x_continuous(limits=c(0, 132)) -->
<!-- ) -->

<!-- sample_size_labs <- c( -->
<!--   `Only pre` = "Only pre (n=705)", -->
<!--   `Only post` = "Only post (n=1158)", -->
<!--   `Both pre and post` = "Both pre and post (n=1940)" -->
<!-- ) -->

<!-- moud %>% -->
<!--   ggplot(aes(x=total.rx.days.exp3, fill=moud.both3div)) +  -->
<!--   geom_histogram(alpha=0.75, color=1) +  -->
<!--   facet_wrap(~moud.both3div, scales = "free_x") + -->
<!--   facetted_pos_scales(x=custom_x) + -->
<!--   labs(x= "Days of MOUD supply", y="Count of pregnancies") + -->
<!--   scale_color_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   scale_fill_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   # guides(fill= guide_legend(title= "Receipt of MOUD relative to delivery")) + -->
<!--   theme_classic() + -->
<!--   theme(legend.box = "vertical", legend.position = "top", text = element_text(size=18))  -->


<!-- moud %>% -->
<!--   ggplot(aes(x=total.rx.days.exp3, fill=moud.both3div)) +  -->
<!--   geom_histogram(alpha=0.75, color = 1) +  -->
<!--   facet_wrap(~moud.both3div, scales = "free_x", labeller = as_labeller(sample_size_labs)) + -->
<!--   facetted_pos_scales(x=custom_x) + -->
<!--   labs(x= "Days of MOUD supply", y="Count of pregnancies", title = "Receipt of MOUD relative to delivery") + -->
<!--   scale_color_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   scale_fill_manual(values=c("black", "darkgray", "tan3")) + -->
<!--   guides(fill= guide_legend(title= "Receipt of MOUD relative to delivery")) + -->
<!--   theme_classic() + -->
<!--   theme(legend.box = "vertical", legend.position = "none", text = element_text(size=18), plot.title = element_text(hjust = 0.5))  -->


<!-- ``` -->

<!-- <br> -->

<!-- # Multivariable Cox PH model -->
<!-- ## Linear continuous terms -->
<!-- ```{r, warning=FALSE, message=NA, include=FALSE} -->

<!-- dd <- datadist(moud) -->
<!-- options(datadist= 'dd') -->
<!-- dd$limits$pain.combine.f[2] <- "No" -->
<!-- dd$limits$alcuse.othersub.f[2] <- "No" -->
<!-- #  -->
<!-- #  -->
<!-- # cph_multi <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.exp3 +  + momage + momraceeth_comb.new + rpl_themes + routedeliv.new + previouspregnanciestotaln + gweeks + -->
<!-- #                                   pain.combine.f + personality.combine.f + alcuse.othersub.f + benzo.anticonv.f + anxiety.f + depression.f + mat_morbidity.f, data=moud, x=TRUE, y=TRUE)  -->

<!-- cph_multi <- cph(Surv(days.to.event, outcome.bin) ~ delivery_year + total.rx.days.exp3 +  momage + routedeliv.new + gweeks +   -->
<!--   pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE)  -->

<!-- ``` -->

<!-- ```{r} -->
<!-- print(summary(robcov(cph_multi, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014, 2019),  vnames = 'labels')) -->
<!-- print(anova(robcov(cph_multi), vnames="labels"), digits=3) -->

<!-- ``` -->

<!-- ```{r, fig.height=6, fig.width=8} -->
<!-- plot(summary(cph_multi, momage=c(25, 35),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014, 2019), vnames = 'labels')) -->
<!-- ``` -->

<!-- ```{r, fig.height=12, fig.width=12} -->
<!-- ggplot(Predict(cph_multi, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical") -->
<!-- ``` -->

<!-- ## With splines -->
<!-- ```{r, include=FALSE, warning=FALSE, message=FALSE} -->

<!-- cph_multi_sp <- cph(Surv(days.to.event, outcome.bin) ~  rcs(delivery_year,3) + total.rx.days.exp3 +  momage + routedeliv.new + gweeks +   -->
<!--   pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE)  -->

<!-- ``` -->

<!-- ```{r} -->
<!-- # print(summary(robcov(cph_multi_sp, cluster = basic$unique_id.x), delivery_year=c(2016, 2020),  momage=c(25, 30), rpl_themes=c(0.5, 0.75), total.rx.days.exp=c(0,1), vnames = 'labels'), digits=3) -->
<!-- # print(anova(robcov(cph_multi_sp), vnames="labels"), digits=3) -->
<!-- # x <- Predict(cph_multi_sp, delivery_year=c(2007:2020)) -->
<!-- # x$yhat_exp <- exp(x$yhat) -->
<!-- # x$lower_exp <- exp(x$lower) -->
<!-- # x$upper_exp <- exp(x$upper) -->

<!-- print(summary(robcov(cph_multi_sp, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014, 2019),  vnames = 'labels')) -->


<!-- print(anova(robcov(cph_multi_sp), vnames="labels"), digits=3) -->

<!-- ``` -->

<!-- ```{r,fig.height=6, fig.width=8} -->
<!-- plot(summary(cph_multi_sp, momage=c(25, 35),  delivery_year=c(2014, 2019),  total.rx.days.exp3=c(1, 31), vnames = 'labels')) -->
<!-- ``` -->

<!-- ```{r, fig.height=12, fig.width=12} -->
<!-- ggplot(Predict(cph_multi_sp, fun=plogis), ylab=expression(hat(P)), sepdiscrete="vertical") -->

<!-- ``` -->

<!-- <br> -->

<!-- ## With splines (on outcome without all payer data) -->
<!-- ```{r, include=FALSE, warning=FALSE, message=FALSE} -->
<!-- moud$outcome.bin2 <- ifelse(moud$outcome2 %in% c("Death", "Overdose"), 1, 0) -->
<!-- cph_multi_sp2 <- cph(Surv(days.to.event, outcome.bin2) ~ rcs(delivery_year,3) + total.rx.days.exp3 +  momage + routedeliv.new + gweeks +   -->
<!--   pain.combine.f + alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f + mat_morbidity.f , data=moud, x=TRUE, y=TRUE)  -->

<!-- ``` -->

<!-- ```{r} -->
<!-- print(summary(robcov(cph_multi_sp2, cluster = moud$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014, 2019),  vnames = 'labels')) -->
<!-- print(anova(robcov(cph_multi_sp2), vnames="labels"), digits=3) -->

<!-- ``` -->

<!-- ### Subgroup analysis (both periods) -->

<!-- ```{r} -->
<!-- moud_both <- moud[moud$moud.both3=="Both pre and post", ] -->
<!-- cph_multi_sp_both <- cph(Surv(days.to.event, outcome.bin) ~ rcs(delivery_year,3) + total.rx.days.exp3 +alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f ,  -->
<!--                          data=moud_both, x=TRUE, y=TRUE)  -->


<!-- print(summary(robcov(cph_multi_sp_both, cluster = moud_both$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014,2019),  vnames = 'labels')) -->
<!-- print(anova(robcov(cph_multi_sp_both), vnames="labels"), digits=3) -->
<!-- ``` -->

<!-- <br> -->


<!-- ### Subgroup analysis (one period) -->

<!-- ```{r} -->
<!-- moud_one <- moud[moud$moud.both3=="Either pre or post", ] -->
<!-- cph_multi_sp_one <- cph(Surv(days.to.event, outcome.bin) ~ rcs(delivery_year,3) + total.rx.days.exp3 +alcuse.othersub.f + benzo.anticonv.f + anx.dep.comb.f ,  -->
<!--                          data=moud_one, x=TRUE, y=TRUE)  -->


<!-- print(summary(robcov(cph_multi_sp_one, cluster = moud_one$unique_id.x),  total.rx.days.exp3=c(1, 31), delivery_year=c(2014,2019),  vnames = 'labels')) -->
<!-- print(anova(robcov(cph_multi_sp_one), vnames="labels"), digits=3) -->
<!-- ``` -->